---
title: "Who should you use breadcrumbs?"
date: "2019-12-20"
output: github_document
---

If you hate being unable to find code and data in old data projects, you may find `breadcrumbs` to be a breath of fresh air. 

The goal for the `breadcrumbs` package is to provide a routine structure to R projects so that you put code and data in the right places and find them later with ease. `breadcrumbs` also encourages you to coordinate your code and data with Makefile(s), which are a wonderful resource for reproducibile workflows.  

### Getting started
To get started using `breadcrumbs`  
1. open a new or existing R project   
2. Load the breadcrumbs package `library(breadcrumbs)`  
2. Execute the `create_project_directories()` function in your console  

You will now see a standard set of subdirectories appear in your project directory. 

#### The subdirectory structure
  `<project>_cloud_dir/` I use the `<project>_cloud_dir/` to save large files like data and image-laden reports that don't belong in GitHub. I recommend you sync this folder to the cloud via Google Drive or Box so that it is backed up off your local machine.   
  ....`reports/` to store finished reports and presentations   
	....`data/` to store processed data  
	........`raw/` to store raw data  
	`R/` to store your functions  
  `munge/` to store scripts that output processed data  
  `analysis/` to store your analysis scripts  

`breadcrumbs` is filled with functions to support the use of these subdirectories. For instance, I have found it useful to save my filepaths to commonly used subdirectories in a dedicated script called "file_paths.R". The function `source_file_paths()` will execute this script, and I include this function at the top of all my other scripts and Rmd files. 

#### Makefiles
`create_project_directories()` has also added a Makefile to your project's home directory. This Makefile can be used to make your entire project reproducible from raw data to finished reports. For more details on how to use Makefiles see the appendix in the Makefile template generated by breadcrumbs. 

### Other useful functions
#### wrap()
In addition to directory structure, `breadcrumbs` includes a number of functions I find useful at the exploratory stage of data projects. For instance `wrap()` will control the width of long ggplot titles and facilitate your writing accurate and informative titles on all your EDA plots without fussing around trying to find the right places to manually insert line breaks. 
```{r message=FALSE, warning=FALSE}
library(breadcrumbs)
library(tidyverse)

ggplot(mtcars, aes(x = wt, y = mpg)) + 
	geom_point() + 
	labs(title = wrap("The wrap function frees me up to write long and informative titles on my plots without having to troubleshoot the placement of linebreaks. This is useful since I often forget what I once found useful about a plot.", 80)) + 
	theme_minimal()
```

#### clean_names()
Another useful function is `clean_names()`. I use `clean_names()` the first time I import raw data to remove capital letters and illegal characters. I'll read in a government dataset below to demonstrate. Note how illegal names like `JURISDICTION NAME` become clean snake_case names like `jurisdiction_name`. This is a great step to take at the start of your project so you don't need to remember esoteric conventions about how variables were names in different datasets. 
```{r message=FALSE, warning=FALSE}
file_nyc_data <- 
    "https://data.cityofnewyork.us/api/views/kku6-nxdu/rows.csv"

# Reading data without clean_names()
read_csv(file_nyc_data) %>% select(1:4)
```

```{r message=FALSE, warning=FALSE}
# Reading data with clean_names()
read_csv(file_nyc_data) %>% select(1:4) %>% clean_names()
```

