---
title: "Why would you find breadcrumbs useful?"
date: "2019-12-20"
output: github_document
---

I hate being unable to find code and data in old data projects. The goal for the `breadcrumbs` package is to provide a routine structure to the subdirectories in your R projects so that you and your collaborators can readily store code and data in appropriate folders, and coordinate that code and data with Makefiles. 

To get started using `breadcrumbs`  
1. open a new or existing R project   
2. Load the breadcrumbs package `library(breadcrumbs)`  
2. Execute the `create_project_directories()` function in your console  

You will now see a standard set of project subdirectories appear in your project directory, and breadcrumbs is filled with functions to support the use of those subdirectories. For instance, I have found it useful to save my filepaths to commonly used subdirectories in a dedicated script called "file_paths.R". The function `source_file_paths()` will execute this script, and I include this function at the top of all my other scripts and Rmd files. `create_project_directories()` has also added a Makefile to your project's home directory. This Makefile can be used to make your entire project reproducible from raw data to finished reports. For more details on how to use Makefiles see the comment appendix in the Makefile generated by breadcrumbs. 

### Other useful functions
In addition to directory structure, `breadcrumbs` includes a number of functions I find useful at the exploratory stage of data projects. For instance `wrap()` will control the width of long ggplot titles and facilitate your writing accurate and informative titles on all your EDA plots without fussing around trying to find the right places to manually insert line breaks. 
```{r message=FALSE, warning=FALSE}
library(breadcrumbs)
library(tidyverse)

ggplot(mtcars, aes(x = wt, y = mpg)) + 
	geom_point() + 
	labs(title = wrap("The wrap function frees me up to write long and informative titles on my plots without having to troubleshoot the placement of linebreaks. This is useful since I often forget what I once found useful about a plot.", 80)) + 
	theme_minimal()
```

Another useful function is `clean_names()`. I use `clean_names()` the first time I import raw data to remove capital letters and illegal characters. 
```{r message=FALSE, warning=FALSE}
file_nyc_data <- 
    "https://data.cityofnewyork.us/api/views/kku6-nxdu/rows.csv"

# Reading data without clean_names()
read_csv(file_nyc_data) %>% select(1:4)
```

```{r message=FALSE, warning=FALSE}
# Reading data with clean_names()
read_csv(file_nyc_data) %>% select(1:4) %>% clean_names()
```

