# Load packages
```{r}
rm(list = ls()) # clear the workspace of all data objects
ls() # list number of data objects in the workspace

# load packages
library(Hmisc)
library(gtools) # for smartbind()
library(psych)
library(corrplot) # for colored cor tables
library(tidyverse) # for %>% 
```

# I. October 2014 Survey 

## Load and clean data
```{r}
# N = 418 for the October 2014 Study
d <- read_csv("03_Oct2014_LDmindset_survey.csv")

dim(d)

d
```

```{r}

# remove useless variables
  d$Intro1 <- NULL
  d$V2 <- NULL          
  d$V3 <- NULL                
  d$V4 <- NULL                
  d$V5 <- NULL                
  d$V6 <- NULL                
  d$V7 <- NULL
  d$V10 <- NULL
  d$LocationLatitude <- NULL
  d$LocationLongitude <- NULL	
  d$LocationAccuracy <- NULL  
  d$Q76 <- NULL
  d$prac <- NULL

# rename variables
  d$Qual_ID <- as.character(d$V1); d$V1 <- NULL
  d$start_time <- d$V8; d$V8 <- NULL
  d$stop_time  <- d$V9; d$V9 <- NULL
  d$howfeel_oe <- d$Q75; d$Q75 <- NULL

# reset variable order
    ## load moveme function
    moveme <- function(data, tomove, where = "last", ba = NULL) {
      temp <- setdiff(names(data), tomove)
      x <- switch(
        where,
        first = data[c(tomove, temp)],
        last = data[c(temp, tomove)],
        before = {
          if (is.null(ba)) stop("must specify ba column")
          if (length(ba) > 1) stop("ba must be a single character string")
          data[append(temp, values = tomove, after = (match(ba, temp)-1))]
        },
        after = {
          if (is.null(ba)) stop("must specify ba column")
          if (length(ba) > 1) stop("ba must be a single character string")
          data[append(temp, values = tomove, after = (match(ba, temp)))]
        })
      x
    }

    # moveme(df, c("b", "c"))
    # moveme(df, c("b", "c"), "first")
    # moveme(df, c("b", "c"), "before", "e") # will move b and c before e.
    # moveme(df, c("b", "c"), "after", "e")

    ## Now move new vars to right places
    colnames(d)
    d <- moveme(d, c("Qual_ID"), "after", "idnum")
```

## Are all student id numbers accurate? 
### How many cases are we starting with? (418)
```{r}
# reclass idnum and initials as character vars
class(d$idnum)
d$idnum <- as.character(d$idnum)
d$initials <- as.character(d$initials)

# record original idnums as idnum_orig
d$idnum_orig <- d$idnum
label(d$idnum_orig) = "idnumber provided by student. Not cleaned or verified against key."
summary(d$idnum_orig)

dim(d)
```
I have 418 cases, but many idnum_orig were mistentered. These were cleaned previously in the file '03_Oct2014_IDcleaning.csv', so I could merge in the cleaned idnumbers based on the Qualtrics idnumbers which are constant across both csv files. 

## How many student-entered idnumbers match the school-provided key? (398/418)
```{r}
# load data of cleaned idnumbers you made last year
cln_ids <- read.csv("03_Oct2014_IDcleaning.csv", head = TRUE)
dim(cln_ids)  # Why are there only 416? One of the missing was CM who took survey later. Who is the other? 
colnames(cln_ids)
cln_ids$Qual_ID <- cln_ids$V1; cln_ids$V1 <- NULL

# merge the cleaned idnumber subset with your Fall2014 data
    ## create a subset for merging
    cln_ids$clnIDdmy <- 1
    cln_ids2 <- subset(cln_ids, select = c(Qual_ID, idnum_clean, clnIDdmy)); dim(cln_ids2)
  
    
    ## Why are there only 416 Ps in your cln_ids data? 
    subset(m, is.na(clnIDdmy), select = c(idnum_orig, initials, start_time))
    >    idnum_orig initials     start_time
    >281      48970       CM 10/29/14 12:39  # took survey after I cleaned data
    >350      18910       ik  10/1/14 16:50  # took survey on time

    ## 18910 logged in twice but took survey only once
    names(cln_ids)
    subset(cln_ids, initials_orig == "ik") # 18910 was redundant in the original data
    subset(cln_ids, idnum_orig == "18910")
    subset(d, idnum_orig == "18910", select = c(initials, start_time, Qual_ID)) # 18910 was redundant in the original data
    subset(d, Qual_ID == "R_d0xTHUIk16iTmoR") # this entry was redundant
    d <- subset(d, Qual_ID != "R_d0xTHUIk16iTmoR"); dim(d) # this entry was redundant and has been dropped

    ## merge the datasets
    m <- merge(d, cln_ids2, by = "Qual_ID", all = TRUE)
    dim(m) 
    colnames(m)

# Let's see the 20 students without idnum_clean
subset(m, is.na(idnum_clean), select = c(idnum_orig, initials)); dim(subset(m, is.na(idnum_clean)))

      ## Tina's 10.22.14 email sheds light on these missing cases. 
      te <- read.csv('03_Oct2014_MissingIdEmail.csv', head = T); dim(te)
      colnames(te)
      ### rename the idnum variable from Tina's email
      te$idnum_orig <- te$idnums.from.survey.10.01.14; te$idnums.from.survey.10.01.14 <- NULL
      ### initials_te & idnum_te are those supplied by Tina's email
      te$initials_te <- te$initials; te$initials <- NULL
      te$idnum_te <- te$Study.ID; te$Study.ID <- NULL # te$idnum_te are accurate idnumbers
     
      # merge data to compare. 
      temp <- merge(m, te, by = "idnum_orig", all = TRUE)
      dim(temp)
      temp2 <- subset(temp, select = c("idnum_clean", "idnum_orig", "idnum_te", "initials_te", "initials", "JP_fill_in"))

      # looks like I'd already incorporated these idnumbers, except 19610
      subset(temp2, !is.na(JP_fill_in))
          # fix the idnum_clean for student __________20160
          m$idnum_clean <- as.numeric(m$idnum_clean)
          m$idnum_clean [m$idnum_orig == "__________20160"] <- 19610
          describe(m$idnum_clean)
          dim(m)
          class(m$idnum_clean)

# clean up data order
names(m)
m <- moveme(m, c("idnum_clean", "idnum_orig"), "before", "Qual_ID")
m$clnIDdmy <- NULL
subset(m, select = c(idnum_clean, idnum_orig, idnum))
m$temp <- ifelse(m$idnum_orig == m$idnum, 1, 0); table(m$temp) # all match
m$temp <- NULL

# Of 417 cases there are 398 cases with recognizable idnumbers
dim(m)
dim(subset(m, !is.na(idnum_clean)))
``` 
So the October 2014 survey I had 417 takers:
As of this chunk 20 are without recognized IDnumbers, 398 have recognized IDnumbers.

## Can you find the idnumbers of those 20 who couldn't be matched? 
### View subset of students without recognized idnumbers
```{r idnumbers of 20 who couldn't be matched}
# subset with unmatched IDs
umid <- subset(m, is.na(idnum_clean)); dim(umid)
names(umid)

# umid$idnum and umid$idnum_orig are redundant
ifelse(umid$idnum == umid$idnum_orig, 1, 0)
    umid$idnum <- NULL

# there are some identifiers in the missing data
sort(umid$idnum_orig)
table(umid$idnum_orig)
subset(umid, select = c(idnum_orig, initials))
```
7 of the 20 entered an unrecognized idnumber
13 of 20 entered the url instead of the website

### Move unrecognized idnum_orig to idnum_clean for cases without an idnum_clean and avoid duplicates.
```{r}
# move unrecognized idnum_origs to idnum_clean for cases without an idnum_clean and avoid duplicates
class(m$idnum_clean)
temp <- subset(m, is.na(idnum_clean), select = c(idnum_clean, Qual_ID, idnum_orig, initials)); temp
temp[order(temp$idnum_orig),] 
summary(m$idnum_clean)# 20 NAs
m$idnum_clean <- as.numeric(m$idnum_clean)

# impose the idnums for 3 cases whose idnumbers were provided by students, but match no school records. 
m$idnum_clean [m$Qual_ID == "R_8B9NZpUsFVBmuBn"] <- 44860 
m$idnum_clean [m$Qual_ID == "R_efAXweKR3ZsQ2Lr"] <- 32500 
m$idnum_clean [m$Qual_ID == "R_9WuHZQgfn2MztRj"] <- 48970 

# These 4 students entered ID numbers that duplicate other ID numbers and don't match the intials entered by the student. 
    # m$idnum_clean [m$Qual_ID == "R_d0xTHUIk16iTmoR"] <- "18910" # would create a duplicate  
    # m$idnum_clean [m$Qual_ID == "R_6YHhjGoTr90bocR"] <- "36390" # would create a duplicate
    # m$idnum_clean [m$Qual_ID == "R_9Ti2V5iEcdza2ZT"] <- "44860" # would create a duplicate     
    # m$idnum_clean [m$Qual_ID == "R_eXvKrSEoz1ZxMJT"] <- "44870" # would create a duplicate

# There are no duplicated IDs under idnum_clean
    sum(duplicated(m$idnum_clean)) 
    m$tempIDdup_check <- duplicated(m$idnum_clean)
    subset(m, tempIDdup_check == TRUE, select = c(idnum_clean))

# There are 401 unique values for idnum_clean
sum(!is.na(m$idnum_clean)) # 401
sort(m$idnum_clean)
```
Of my 418 participants, 401 have unique idnumbers in idnum_clean. 
398 of the 401 have idnumbers that are matched with school records. The difference of 3 represents people who provided a clean idnumber that matched no records, or duplicated another idnumber while not matching provided student initials. 

## I have to give these students temporary idnumbers to enable later merging. 
```{r temp id numbers for NA idnum_clean}
# Find the Qualtrics IDs for those without clean idnumbers
newIDs <- subset(m, is.na(idnum_clean)); dim(newIDs)
newIDs$Qual_ID

# Now force temporary IDs upon them
class(m$idnum_clean)
m$idnum_clean [m$Qual_ID == "R_03u2CMOUDp8XHhP"] <- 99999001
m$idnum_clean [m$Qual_ID == "R_086J8YtIDQw2LXf"] <- 99999002
m$idnum_clean [m$Qual_ID == "R_1AmbXuyNDFZiDSR"] <- 99999003 
m$idnum_clean [m$Qual_ID == "R_3DZB8hBNbuaajSl"] <- 99999004
m$idnum_clean [m$Qual_ID == "R_4UA4eFzJnipXn93"] <- 99999005 
m$idnum_clean [m$Qual_ID == "R_5oPdfJEq2Ej39zf"] <- 99999006 
m$idnum_clean [m$Qual_ID == "R_6YHhjGoTr90bocR"] <- 99999007 
m$idnum_clean [m$Qual_ID == "R_73NSYm22eisYcNn"] <- 99999008
m$idnum_clean [m$Qual_ID == "R_9Ti2V5iEcdza2ZT"] <- 99999009 
m$idnum_clean [m$Qual_ID == "R_ahsO23AVLj4UC9v"] <- 99999010 
m$idnum_clean [m$Qual_ID == "R_b2BdU6zofpd3CXb"] <- 99999011 
m$idnum_clean [m$Qual_ID == "R_b2YyU92nsc89Uhf"] <- 99999012
m$idnum_clean [m$Qual_ID == "R_eeXCD4BAhoWXb8x"] <- 99999013 
m$idnum_clean [m$Qual_ID == "R_emS99MXNLIshgBn"] <- 99999014 
m$idnum_clean [m$Qual_ID == "R_eOFk15lLLSoxiND"] <- 99999015
m$idnum_clean [m$Qual_ID == "R_eXvKrSEoz1ZxMJT"] <- 99999016

class(m$idnum_clean)

sum(is.na(m$idnum_clean)) # No NAs; all Ps have IDs now
```

## Is there any remaining way to ID these 16 Ps with temporary IDnumbers? 
```{r}
# I have their intials
subset(newIDs, select = c(initials, idnum_orig))
names(d)

# all were under 18
subset(newIDs, select = c("minor", "consent", "assent")) 

# Are there no unclaimed initials that match? 

# Do later surveys or grade records offer indication of who they might be?
```
__ detective work remains for 16 Ps with temporary idnumbers in chunk above!!!

## tag the data frame from Fall 2014 with _f14
```{r}
library(dplyr)
names(m)[2:length(names(m))] <- paste0(names(m)[2:length(names(m))], "_f14")
```

# II. 2013-14 Prior Performance and Demographic Data

## Demographic and Progress Testing Data
```{r}
# main data 417 cases with 16 temp idnumbers I added to those without idnumbers. temp IDs begin 999. 
dim(m); sum(is.na(m$idnum_clean))

# setwd for data to merge
setwd("~/Box Sync/LD/10_LD_data_merging/04_2013-14_LDstudentsData")

# Demographic data
dem <- read.csv('Student_Demographic_FINAL.csv', head = T)
    dim(dem); names(dem)   # 448 5
    dem$idnum_clean <- as.numeric(dem$StudentID); dem$StudentID <- NULL
    sum(is.na(dem$idnum_clean)) # 0 NAs
    sum(duplicated(dem$idnum_clean)) # 0 duplicates

# Progress testing data
prog <- read.csv('Student_ProgressTesting_FINAL.csv', head = T)
    dim(prog); names(prog) # 445  23
    prog$idnum_clean <- as.numeric(prog$StudentID) 
      prog$StudentID <- NULL
    sum(is.na(prog$idnum_clean)) # 0 NAs
    sum(duplicated(prog$idnum_clean)) # 1 duplicate 
        # Why is there a duplicate? 
        # 33420 has two sets of scores, take the more recent WISC drop the older WISC, average GORTs
        temp33420 <- subset(prog, idnum_clean == 33420)
        # write.csv(temp33420, "temp33420.csv") # this file was then processed where WISC recent scores and GORT average scores were used because WISC was given on different dates and GORT was given twice on same date, and then merged back into the file Student_ProgressTestingFINAL to create Student_ProgressTesting_FIXED33420.
        # The double score for GORT on one day is super strange and I'll ask Tina about it later!!!
        prog <- read.csv("Student_ProgressTesting_FIXED33420.csv", head = T)
        dim(prog) # 444 23
        prog$idnum_clean <- as.numeric(prog$StudentID) 
        prog$StudentID <- NULL
```
So I have Progress Testing on 444 students and Demo data on 448. 

```{r}
# begin merging
  # temp IDs for testing the merge
  m$test_id_m <- m$idnum_clean
  dem$test_id_dem <- dem$idnum_clean
  names(dem)
  dem <- moveme(dem, c("idnum_clean"), "before", "Birthdate") 
  # tag the dem data 
  names(dem)[2:length(names(dem))] <- paste0(names(dem)[2:length(names(dem))], "_1314")
  names(dem)

  # merge
  m2 <- merge(m, dem, by = "idnum_clean", all = TRUE)
  dim(m2) # 466
      # test merge
      m2$test <- m2$test_id_m - m2$test_id_dem; table(m2$test) # all zeroes confirms clean merge
      # drop testing vars
      m2$test <- NULL; m2$test_id_m <- NULL;  m2$test_id_dem <- NULL 
  # temp IDs for testing the merge
  m2$test_id_m <- m2$idnum_clean
  prog$test_id_prog <- prog$idnum_clean
  
  # tag the prog data 
  names(prog)
  prog <- moveme(prog, c("idnum_clean"), "before", "Pattern") 
  names(prog)[2:length(names(prog))] <- paste0(names(prog)[2:length(names(prog))], "_1314")
  names(prog)

  # merge
  m3 <- merge(m2, prog, by = "idnum_clean", all = TRUE)
  dim(m3) # 466 
      # test merge
      m3$test <- m3$test_id_m - m3$test_id_prog; table(m3$test) # all zeroes confirms clean merge
      m3$test <- NULL; m3$test_id_m <- NULL; m3$test_id_prog <- NULL;
  # confirm no NAs or duplicates for idnum_clean
  sum(is.na(m3$idnum_clean))
  sum(duplicated(m3$idnum_clean)) # 0

  # overwrite m with m3
  m <- m3
```

```{r calculate composite variables for gort and WISC IQ}
m$gort_1314 <- apply(X = m[,c("GORT.Reading.Accur_1314",  "GORT.Reading.Comp_1314",  "GORT.Reading.Rate_1314")],
                MARGIN =1, FUN = mean, na.rm = T)
  m <- moveme(m, c("gort_1314"),"before", "GORT.Reading.Accur_1314")
  label(m$gort1314) = "mean GORT score from 2013-14 data" 
  describe(m$gort1314)
  hist(m$gort1314)

m$WISC_1314 <- apply(X = m[,c("WISC.WorkingMemory_1314","WISC.VerbalComp_1314","WISC.PerceptualReasoning_1314","WISC.ProcessingSpeed_1314")],
                MARGIN =1, FUN = mean, na.rm = T)
  m <- moveme(m, c("WISC_1314"),"after", "WISC_TestDAte_1314")
  label(m$WISC_1314) = "mean WISC IQ score from 2013-14 data" 
  describe(m$WISC_1314)
  hist(m$WISC_1314)
```

### Sort by Pattern (~text-to-column)
This chunk is to text-to-column the Pattern data and then to merge it back in with the main dataset. 
```{r Sort by pattern}
library(splitstackshape)
dim(m)
t <- subset(m, select = c(idnum_clean, Pattern_1314)); dim(t) # 466   2
t$idnum_clean # bunch of NAs at end. Checking above df's to see why.
t2 <- cSplit(t, 'Pattern_1314', sep=", ", type.convert=FALSE)
names(t2) # no one had more than 6 Patterns

i <- 1
t2$Pattern1_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern1_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern1_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern1_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern1_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern1_1314 [t2$Pattern_1314_6 == i] <- i

i <- 2
t2$Pattern2_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern2_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern2_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern2_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern2_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern2_1314 [t2$Pattern_1314_6 == i] <- i

i <- 3
t2$Pattern3_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern3_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern3_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern3_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern3_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern3_1314 [t2$Pattern_1314_6 == i] <- i

i <- 4
t2$Pattern4_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern4_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern4_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern4_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern4_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern4_1314 [t2$Pattern_1314_6 == i] <- i

i <- 5
t2$Pattern5_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern5_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern5_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern5_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern5_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern5_1314 [t2$Pattern_1314_6 == i] <- i

i <- 6
t2$Pattern6_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern6_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern6_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern6_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern6_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern6_1314 [t2$Pattern_1314_6 == i] <- i

i <- 7
t2$Pattern7_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern7_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern7_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern7_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern7_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern7_1314 [t2$Pattern_1314_6 == i] <- i

i <- 8
t2$Pattern8_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern8_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern8_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern8_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern8_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern8_1314 [t2$Pattern_1314_6 == i] <- i

i <- 9
t2$Pattern9_1314 [t2$Pattern_1314_1 == i] <- i
t2$Pattern9_1314 [t2$Pattern_1314_2 == i] <- i
t2$Pattern9_1314 [t2$Pattern_1314_3 == i] <- i
t2$Pattern9_1314 [t2$Pattern_1314_4 == i] <- i
t2$Pattern9_1314 [t2$Pattern_1314_5 == i] <- i
t2$Pattern9_1314 [t2$Pattern_1314_6 == i] <- i

head(t2)
describe(t2$Pattern1_1314)
describe(t2$Pattern2_1314)
describe(t2$Pattern3_1314)
describe(t2$Pattern4_1314)
describe(t2$Pattern5_1314)
describe(t2$Pattern6_1314)
describe(t2$Pattern7_1314)
describe(t2$Pattern8_1314)
describe(t2$Pattern9_1314)

# Remove variables you don't want merged back in with the main dataset
t2$Pattern_1314_1 <- NULL
t2$Pattern_1314_2 <- NULL
t2$Pattern_1314_3 <- NULL
t2$Pattern_1314_4 <- NULL
t2$Pattern_1314_5 <- NULL
t2$Pattern_1314_6 <- NULL
names(t2)

class(t2)
t2 <- data.frame(t2)
dim(t2); colnames(t2); class(t2)
t2$idnum_clean


# merge with m, the survey data with complete idnumbers
m$testidm <- m$idnum_clean
t2$testidt <- t2$idnum_clean
m <- merge(m, t2, by = "idnum_clean", all = TRUE); dim(m)
  # test merger
  m$test <- m$testidm - t2$testidt; table(m$test)
  m$test <- NULL; m$testidm <- NULL; t2$testidt <- NULL
  # check the merger
  length(unique(m$idnum_clean)) # 466 
  sum(duplicated(m$idnum_clean)) # 0
  sum(is.na(m$idnum_clean)) # 0

dim(m) # 466 109
names(m)

# drop extra vars
m$testidt <- NULL
```

## Grades data need to be recast as wide. 
```{r convert grades from long to wide}
# Demographic and Progress Testing Files look like wide data

# set wd to 2013-14 grades folder
setwd("~/Box Sync/LD/10_LD_data_merging/04_2013-14_LDstudentsData")

# read in grades with na.strings = "" so that NAs appear in blank spaces. 
gr <- read.csv("Student_ClassroomGrades_FINAL.csv", na.strings = "", head = T)
dim(gr); names(gr)
head(gr)
class(gr$Course)

sum(is.na(gr$Course)) # 164 course performance grades have no course name associated with them. Might these be final grades for the term? FUP!!!

# Prepare grades data for merging
# paste course identyifing data into one variable course_yr_term
    gr$Course2 <- gsub("[ ]", "_", gr$Course)
    gr$course_yr_trm <- paste(gr$Course2, "_1314t", gr$Semester_Term, sep="")
# rename performance variable
gr$score <- gr$Grade; gr$Grade <- NULL

# Convert data from long to wide format using cast()
  # create subset that drops nameless courses from data
  cp <- subset(gr, !is.na(Course))
  dim(cp)
  sum(is.na(cp$StudentID)) # 0 NAs

  # view key variables
  describe(cp$StudentID); length(unique(cp$StudentID)) # 349 unique IDs
  
  summary(cp$course_yr_trm)
  describe(cp$score)
  
  cp$idnum_clean <- cp$StudentID
  
  # create a subset of the variables you will transform to a wide dataset
  cp2 <- subset(cp, select = c(idnum_clean, course_yr_trm, score))
  dim(cp2)
  head(cp2)

# load packages
library(reshape2)
w <- dcast(cp2, idnum_clean ~ course_yr_trm,
           value.var = "score",
           fun.aggregate = mean, 
           na.rm = TRUE)
      # what is the effect of na.rm=T here???
      # what is the effect of mean in fun.aggregate???
      # how can I confirm if any data were lost???

dim(w) 
  # 349 unique idnumber matches the number of unique IDs in describe(cp$StudentID)
  # 167 + 1 (ID col) columns matches describe(cp$course_yr_trm) values for unique course_yr_trm values
# test <- write.csv(w, 'test.csv') 
```

## tag and merge your 2013-14 prior performance data with the fall 2014 data
Did not need to tag grading variables as this had already occurred
```{rmerge}
names(w)
# Merge your wide 1314 grades data with the fall 2014 data
dim(w) 
dim(m)

# test vars for merger
w$testw <- w$idnum_clean
m$testm <- m$idnum_clean

m2 <- merge(m, w, by = "idnum_clean", all = TRUE); dim(m2) # 466
m2$test <- m2$testw-m2$testm; table(m2$test)
m2$test <- NULL; m2$testw <- NULL; m2$testm <- NULL

# merger looks good, so overwrite
m <- m2
names(m)
```
"w" is the wide data. 
__ Looks good but have RAs spot check 10 cases each.

```{r}
# How many took the Fall 2014 survey? 
dim(m) # [1] 466 279
sum(!is.na(m$Qual_ID)) # 417 took the survey is consistent with earlier figures. 
sum(duplicated(m$Qual_ID)) # these 48 duplicated Qual_IDs are the NAs who I have data for, but did not take survey
  m$temp_chk <- duplicated(m$Qual_ID)
  subset(m, temp_chk == "TRUE", select = c(Qual_ID))

    # drop extra var 
    m$temp_chk <- NULL
```
At this point I have merged the Fall 2014 data and the grades data from the prior year. 
There are 466 people in my data. 
417 survey takers and 49 additional cases for whom I just have data. (10/21/15)

# III. Spring 2015 Mindset intervention data

## Mindset Data from Spring 2015
Reduce dataframe size by dropping useless variables. 
```{r drop timing variables from survey for now}
setwd("~/Box Sync/LD/10_LD_data_merging")
ms <- read.csv("05_LD_Apr2015_master.csv", na.strings = "", head = T)
dim(ms) # [1] 414 638
sum(is.na(ms$idnum))
```

```{r drop excess variables and relabel vars}
# relabel vars
ms$Qual_ID <- as.character(ms$V1); ms$V1 <- NULL
ms$EndApr2015 <- ms$V9; ms$V9 <- NULL
ms$StartApr2015 <- ms$V8; ms$V8 <- NULL
ms$read_auton <- ms$Q336; ms$Q336 <- NULL
ms$read_ctrl <- ms$Q335; ms$Q335 <- NULL
ms$read_help <- ms$Q334; ms$Q334 <- NULL
ms$read_imp <- ms$Q333; ms$Q333 <- NULL
ms$read_days <- ms$Q332; ms$Q332 <- NULL
ms$rc_see <- ms$Q187_4; ms$Q187_4 <- NULL
ms$rc_apsee <- ms$Q187_3; ms$Q187_3 <- NULL
ms$rc_rass <- ms$Q187_2; ms$Q187_2 <- NULL
ms$rc_ap <- ms$Q187_1; ms$Q187_1 <- NULL
ms$cnt_auton <- ms$Q170; ms$Q170 <- NULL
ms$cnt_ctrl <- ms$Q169; ms$Q169 <- NULL
ms$vid_prob_oe <- ms$Q161; ms$Q161 <- NULL
ms$vid_play <- ms$Q160; ms$Q160 <- NULL
ms$cond_ms <- ms$DO.BR.FL_9; ms$DO.BR.FL_9 <- NULL
ms$cnt_use <- ms$CNT_use; ms$CNT_use <- NULL
ms$cnt_imp <- ms$CNT_imp; ms$CNT_imp <- NULL
ms$cnt_descr <- ms$CNT_descr; ms$CNT_descr <- NULL
ms$cnt_burden <- ms$CNT_burden; ms$CNT_burden <- NULL
ms$bel <- ms$Q457; ms$Q457 <- NULL

# variables to drop flagged in excel and transpose pasted below
ms2 <- subset(ms, select = -c(assent,   forebr, 	grow_out, 	intr1, 	kayla, 	new.link, 	Q1, 	Q13, 	Q133t_1, 	Q133t_2, 	Q133t_3, 	Q133t_4, 	Q139t_1, 	Q139t_2, 	Q139t_3, 	Q139t_4, 	Q13t_1, 	Q13t_2, 	Q13t_3, 	Q13t_4, 	Q15, 	Q152, 	Q153_1, 	Q153_2, 	Q153_3, 	Q153_4, 	Q155_1, 	Q155_2, 	Q155_3, 	Q155_4, 	Q156, 	Q156t_1, 	Q156t_2, 	Q156t_3, 	Q156t_4, 	Q158, 	Q159_1, 	Q159_2, 	Q159_3, 	Q159_4, 	Q161_1, 	Q161_2, 	Q161_3, 	Q161_4, 	Q162, 	Q163_1, 	Q163_2, 	Q163_3, 	Q163_4, 	Q164, 	Q165_1, 	Q165_2, 	Q165_3, 	Q165_4, 	Q166, 	Q167_1, 	Q167_2, 	Q167_3, 	Q167_4, 	Q168, 	q168t_1, 	q168t_2, 	q168t_3, 	q168t_4, 	Q17, 	Q171_1, 	Q171_2, 	Q171_3, 	Q171_4, 	Q172, 	q172t_1, 	q172t_2, 	q172t_3, 	q172t_4, 	Q173_1, 	Q173_2, 	Q173_3, 	Q173_4, 	Q174_1, 	Q174_2, 	Q174_3, 	Q174_4, 	Q174, 	Q174t_1, 	Q174t_2, 	Q174t_3, 	Q174t_4, 	Q175_1, 	Q175_2, 	Q175_3, 	Q175_4, 	Q176, 	Q177_1, 	Q177_2, 	Q177_3, 	Q177_4, 	Q178, 	Q179_1, 	Q179_2, 	Q179_3, 	Q179_4, 	Q181_1, 	Q181_2, 	Q181_3, 	Q181_4, 	Q182, 	Q183_1, 	Q183_2, 	Q183_3, 	Q183_4, 	Q184, 	q184t_1, 	q184t_2, 	q184t_3, 	q184t_4, 	Q186_1, 	Q186_2, 	Q186_3, 	Q186_4, 	Q186, 	q186t_1, 	q186t_2, 	q186t_3, 	q186t_4, 	Q188_1, 	Q188_2, 	Q188_3, 	Q188_4, 	Q188, 	Q189_1, 	Q189_2, 	Q189_3, 	Q189_4, 	Q189, 	Q19, 	Q190_1, 	Q190_2, 	Q190_3, 	Q190_4, 	Q190, 	q190t_1, 	q190t_2, 	q190t_3, 	q190t_4, 	Q192_1, 	Q192_2, 	Q192_3, 	Q192_4, 	Q192, 	Q193_1, 	Q193_2, 	Q193_3, 	Q193_4, 	Q194_1, 	Q194_2, 	Q194_3, 	Q194_4, 	Q195_1, 	Q195_2, 	Q195_3, 	Q195_4, 	Q198_1, 	Q198_2, 	Q198_3, 	Q198_4, 	Q199_1, 	Q199_2, 	Q199_3, 	Q199_4, 	q1t_1, 	q1t_2, 	q1t_3, 	q1t_4, 	Q200_1, 	Q200_2, 	Q200_3, 	Q200_4, 	Q201_1, 	Q201_2, 	Q201_3, 	Q201_4, 	Q203_1, 	Q203_2, 	Q203_3, 	Q203_4, 	Q204_1, 	Q204_2, 	Q204_3, 	Q204_4, 	Q205_1, 	Q205_2, 	Q205_3, 	Q205_4, 	Q207_1, 	Q207_2, 	Q207_3, 	Q207_4, 	Q208_1, 	Q208_2, 	Q208_3, 	Q208_4, 	Q209_1, 	Q209_2, 	Q209_3, 	Q209_4, 	Q210_1, 	Q210_2, 	Q210_3, 	Q210_4, 	Q211_1, 	Q211_2, 	Q211_3, 	Q211_4, 	Q212_1, 	Q212_2, 	Q212_3, 	Q212_4, 	Q213_1, 	Q213_2, 	Q213_3, 	Q213_4, 	Q214_1, 	Q214_2, 	Q214_3, 	Q214_4, 	Q215_1, 	Q215_2, 	Q215_3, 	Q215_4, 	Q216_1, 	Q216_2, 	Q216_3, 	Q216_4, 	Q217_1, 	Q217_2, 	Q217_3, 	Q217_4, 	Q217, 	Q219_1, 	Q219_2, 	Q219_3, 	Q219_4, 	Q221_1, 	Q221_2, 	Q221_3, 	Q221_4, 	Q222_1, 	Q222_2, 	Q222_3, 	Q222_4, 	Q223_1, 	Q223_2, 	Q223_3, 	Q223_4, 	Q224_1, 	Q224_2, 	Q224_3, 	Q224_4, 	Q225_1, 	Q225_2, 	Q225_3, 	Q225_4, 	Q226_1, 	Q226_2, 	Q226_3, 	Q226_4, 	Q227_1, 	Q227_2, 	Q227_3, 	Q227_4, 	Q228_1, 	Q228_2, 	Q228_3, 	Q228_4, 	Q229_1, 	Q229_2, 	Q229_3, 	Q229_4, 	Q23, 	Q230_1, 	Q230_2, 	Q230_3, 	Q230_4, 	Q231_1, 	Q231_2, 	Q231_3, 	Q231_4, 	Q232_1, 	Q232_2, 	Q232_3, 	Q232_4, 	Q233_1, 	Q233_2, 	Q233_3, 	Q233_4, 	Q234_1, 	Q234_2, 	Q234_3, 	Q234_4, 	Q235_1, 	Q235_2, 	Q235_3, 	Q235_4, 	Q236_1, 	Q236_2, 	Q236_3, 	Q236_4, 	Q237_1, 	Q237_2, 	Q237_3, 	Q237_4, 	Q238_1, 	Q238_2, 	Q238_3, 	Q238_4, 	Q239_1, 	Q239_2, 	Q239_3, 	Q239_4, 	Q240_1, 	Q240_2, 	Q240_3, 	Q240_4, 	Q241_1, 	Q241_2, 	Q241_3, 	Q241_4, 	Q242_1, 	Q242_2, 	Q242_3, 	Q242_4, 	Q243_1, 	Q243_2, 	Q243_3, 	Q243_4, 	Q244_1, 	Q244_2, 	Q244_3, 	Q244_4, 	Q245_1, 	Q245_2, 	Q245_3, 	Q245_4, 	Q246_1, 	Q246_2, 	Q246_3, 	Q246_4, 	Q247_1, 	Q247_2, 	Q247_3, 	Q247_4, 	Q248_1, 	Q248_2, 	Q248_3, 	Q248_4, 	Q249_1, 	Q249_2, 	Q249_3, 	Q249_4, 	Q25, 	Q250_1, 	Q250_2, 	Q250_3, 	Q250_4, 	Q251_1, 	Q251_2, 	Q251_3, 	Q251_4, 	Q252_1, 	Q252_2, 	Q252_3, 	Q252_4, 	Q253_1, 	Q253_2, 	Q253_3, 	Q253_4, 	Q254_1, 	Q254_2, 	Q254_3, 	Q254_4, 	Q255_1, 	Q255_2, 	Q255_3, 	Q255_4, 	Q256_1, 	Q256_2, 	Q256_3, 	Q256_4, 	Q257_1, 	Q257_2, 	Q257_3, 	Q257_4, 	Q258_1, 	Q258_2, 	Q258_3, 	Q258_4, 	Q259_1, 	Q259_2, 	Q259_3, 	Q259_4, 	Q265_1, 	Q265_2, 	Q265_3, 	Q265_4, 	Q266_1, 	Q266_2, 	Q266_3, 	Q266_4, 	Q267_1, 	Q267_2, 	Q267_3, 	Q267_4, 	Q268_1, 	Q268_2, 	Q268_3, 	Q268_4, 	Q269_1, 	Q269_2, 	Q269_3, 	Q269_4, 	Q270_1, 	Q270_2, 	Q270_3, 	Q270_4, 	Q271_1, 	Q271_2, 	Q271_3, 	Q271_4, 	Q272_1, 	Q272_2, 	Q272_3, 	Q272_4, 	Q273_1, 	Q273_2, 	Q273_3, 	Q273_4, 	Q274_1, 	Q274_2, 	Q274_3, 	Q274_4, 	Q275_1, 	Q275_2, 	Q275_3, 	Q275_4, 	Q276_1, 	Q276_2, 	Q276_3, 	Q276_4, 	Q278_1, 	Q278_2, 	Q278_3))
# still dropping variables
ms3 <- subset(ms2, select = -c(V3, V4, V5, V6, V7, V10, Q278_4, 	Q280_1, 	Q280_2, 	Q280_3, 	Q280_4, 	Q281_1, 	Q281_2, 	Q281_3, 	Q281_4, 	Q282_1, 	Q282_2, 	Q282_3, 	Q282_4, 	Q283_1, 	Q283_2, 	Q283_3, 	Q283_4, 	Q284_1, 	Q284_2, 	Q284_3, 	Q284_4, 	Q286_1, 	Q286_2, 	Q286_3, 	Q286_4, 	Q287_1, 	Q287_2, 	Q287_3, 	Q287_4, 	Q288_1, 	Q288_2, 	Q288_3, 	Q288_4, 	Q29, 	Q291_1, 	Q291_2, 	Q291_3, 	Q291_4, 	Q292_1, 	Q292_2, 	Q292_3, 	Q292_4, 	Q293_1, 	Q293_2, 	Q293_3, 	Q293_4, 	Q294_1, 	Q294_2, 	Q294_3, 	Q294_4, 	Q295_1, 	Q295_2, 	Q295_3, 	Q295_4, 	Q3, 	Q31, 	Q33, 	Q35, 	Q39, 	Q41, 	Q43, 	Q45, 	q457t_1, 	q457t_2, 	q457t_3, 	q457t_4, 	Q47, 	Q49, 	Q5, 	q5t_1, 	q5t_2, 	q5t_3, 	q5t_4, 	Q7, 	Q9, 	s2.ben, 	s3nt, 	SSins2, 	stratpk_1, 	stratpk_2, 	stratpk_3, 	V2, LocationLatitude,LocationLongitude,LocationAccuracy, Q216, Q218, Q139, Q133, IDS, saf_ck, Q195, Intro1, X4lobe, Q154))

dim(ms3)
ms$Q457
names(ms3)

# write.csv(ms3, "orderms3.csv") and sort into 05_impose order.csv using vlookup. Then open 05_impose order.csv and copy/paste in the order from the console

o <- read.csv('05_impose order.csv', head = T)
as.character(o$vars)
ms4 <- ms3[,c("idnum", "Qual_ID", "cond_ms", "StartApr2015", "EndApr2015", "initials", "current_gr", "start_gr", "dob", "stratxp_oe", "vid", "vid_play", "vid_prob_oe", "image", "image_TEXT", "RC_ins", "ITI_amount", "ITI_change", "ITI_basic", "SM1_r", "SM2", "SM_scav", "tr_best", "tr_control", "tr_fair", "freq", "cnt_descr", "cnt_use", "cnt_imp", "cnt_burden", "cnt_ctrl", "cnt_auton", "rc_ap", "rc_rass", "rc_apsee", "rc_see", "MDR_ex", "read_days", "read_imp", "read_help", "read_ctrl", "read_auton", "org_descr", "org_freq", "org_imp", "org_waste", "org_ctrl", "org_auton", "org_fut", "sara_ex", "sara_days", "sara_helps", "sara_unhlp", "sara_ctrl", "sara_aut", "str_descr", "str_freq", "str_imp", "strl_burd", "str_ctrl", "str_auton", "SS_B1", "SS_B2", "SS_B3", "SS_F4", "SS_F7", "SS_A8", "SS_A9", "SS_A10", "SS_F6", "SS_A11", "SS_A12", "B_E1", "B_E2", "B_A1", "B_SP1", "B_SP2", "B_SP3", "B_SM4", "B_SM5", "B_SM6", "B_SM7r", "hrs_work", "anyelse", "wrong.id")]

names(ms4)
dim(ms4)
ms4$Q154
# overwrite ms4 into m
ms <- ms4
# moveme(df, c("b", "c"), "before", "e") 

# reclass vars
ms$dob <- as.character(ms$dob)

```
414 cases in the Mindset Intervention df (ms)

## cleanup the idnumbers in the Spring 2015 intervention data

### write in temp IDs for idnums that are NAs to avoid merging problems
#### Drop cases that provided no data at all
```{r}
# View cases with no idnum
subset(ms, is.na(idnum)) # they have no data and all but 1 were logged in for less than two minutes. One was logged in for over an hour. 

# drop these cases with no idnums
dim(ms) # 414 before the drop
ms2 <- subset(ms, !is.na(idnum))
dim(ms2) # 398 after dropping is.na(idnum)
ms <- ms2 # overwrite the clean df upon the older df
```
398 cases in ms after dropping empty rows. 

First create a df of intials and demographic identifiers
```{r create a key of inittials and other identifying records called id}
setwd("~/Box Sync/LD/10_LD_data_merging")
init <- read.csv("03_Oct2014_Student_Initials_ID.csv", head = T)
names(init)
dim(init)
init$idnum_clean <- init$Student_ID_Transform; init$Student_ID_Transform <- NULL
init$initials <- init$Initials_CALC; init$Initials_CALC <- NULL

setwd("~/Box Sync/LD/10_LD_data_merging/04_2013-14_LDstudentsData")
dem <- read.csv("Student_Demographic_FINAL.csv", head = T)
dim(dem)
names(dem)
dem$idnum_clean <- dem$StudentID; dem$StudentID <- NULL


# merge your data
id <- merge(dem, init, by = "idnum_clean", all = TRUE)
dim(id)
names(id)
id$idnum_clean <- as.character(id$idnum_clean)
id$Birthdate <- as.character(id$Birthdate) 




# write.csv(id, "id_init_demo.csv")
```

## Populate ms$idnum_clean with trustworthy idnumbers
### create ms$idnum_clean variable
```{r mindset data cleanup}
# convert idnum to a character to deal with text in responses
ms$idnum <- as.character(ms$idnum) # this is the variable you will be cleaning
ms$idnum_orig <- ms$idnum # for reference only

# create idnum_clean to fill in with verified idnumbers
ms$idnum_clean <- NA
ms$idnum_clean <- as.character(ms$idnum_clean)
```

### Update 4 digit idnums with a zero at end
```{r}
# how many idnums < 5 characters
dim(subset(ms, nchar(idnum) < 5)) # 38 < 5 digit idnumber

# 36 4-digit idnumbers
dim(subset(ms, nchar(idnum) == 4, select = idnum))

# update 4-digit idnums with a zero on end 
    # testing conversion
    ms$idnum_temp <- ifelse(nchar(ms$idnum) == 4, 
                            paste0(ms$idnum, "0"), 
                            NA)
    t <- subset(ms, !is.na(idnum_temp), select = c(idnum_temp, idnum, initials, dob)); t; dim(t)
          # conversion looks good
    # check against id key
    m_t <- merge(t, id, by.x = "idnum_temp", by.y = "idnum_clean"); m_t
    subset(m_t, select = c(dob, Birthdate)) # birthdates align
    subset(m_t, select = c(initials.x, initials.y)) # initials align
    
    # update ms$idnum_clean with verified numbers in m_t
    ## prep m_t for merging
    sum(duplicated(m_t$idnum_temp)) # no dups
    ## rename variables for merging
    m_t$idnum_verified <- m_t$idnum_temp
        ## reduce m_t to key variables: idnum and idnum_verified
        m_t <- subset(m_t, select = c(idnum, idnum_verified)); dim(m_t) # 33
    # check key var for duplicates
    sum(duplicated(m_t$idnum)) # 0 dups
    sum(duplicated(ms$idnum)) # 6 dups
      # will any of the 6 dups in ms effect my merger? 
      table(subset(ms, duplicated(idnum) == "TRUE") %in% m_t$idnum) # no effect
    # create idnum_ck vars
      ms$idnum_ck1 <- as.numeric(ms$idnum)
      m_t$idnum_ck2 <- as.numeric(m_t$idnum)
    # update ms$idnum_clean with verified numbers
    ms_mergeCK <- merge(ms, m_t, by = "idnum", all = T)
    dim(ms); dim(m_t)
    dim(ms_mergeCK) # dim looks good
    subset(ms_mergeCK, select = c(idnum, idnum_verified))
    # check merger
    table(ms_mergeCK$idcheck <- ms_mergeCK$idnum_ck2 - ms_mergeCK$idnum_ck1)
    # overwrite dataframes
    ms <- ms_mergeCK; ms_mergeCK <- NULL
    # update master dataframe (ms)
    ms$idnum_clean <- ifelse(!is.na(ms$idnum_verified), 
                            ms$idnum_verified,
                            ms$idnum_clean)
    dim(subset(ms, !is.na(idnum_clean))) # confirms 33 idnum_clean values added

# compare 4 digit idnum and idnum_temp
subset(ms, !is.na(idnum_temp), select = c(idnum_temp, idnum))
dim(subset(ms, !is.na(idnum_temp), select = c(idnum_temp, idnum)))
  # overwrite idnums of 4 digits
    describe(ms$idnum)
    ms$idnum <- ifelse(nchar(ms$idnum) == 4, ms$idnum_temp, ms$idnum)
# how many idnums < 5 characters
dim(subset(ms, nchar(idnum) < 5)) # 2
```
33 clean idnums in ms now.

### find and correct students who entered last name in idnum
Just updating idnum and idnum_clean with this chunk
```{r find and correct students who entered last name in idnum}
# view idnums corrupted with names
sort(ms$idnum)
< "boiko19240" "Case78830"  "grimmer0"   "Judge20880" "kyle36870"  "smith25910"

# check if initials make sense for name and idnumber and relabel idnum_clean accordingly
    subset(id, idnum_clean == 19240, select = c(initials, Birthdate))
    subset(ms, idnum == "boiko19240", select = c(initials, dob))
    ms$idnum [ms$idnum == "boiko19240"] <- 19240
    ms$idnum_clean <- ifelse(ms$idnum == "19240", ms$idnum, ms$idnum_clean)
    
    subset(id, idnum_clean == 78830, select = c(initials)) # not found 
    subset(ms, idnum == 78830, select = c(initials, dob)) # not found, but no reason to doubt b/c no one else used this id in the intervention data. 
    ms$idnum [ms$idnum == "Case78830"] <- 78830
    ms$idnum_clean <- ifelse(ms$idnum == "78830", ms$idnum, ms$idnum_clean)

    subset(id, idnum_clean == 20880, select = c(initials, Birthdate))
    subset(ms, idnum == "Judge20880", select = c(initials, dob))
    ms$idnum [ms$idnum == "Judge20880"] <- 20880
    ms$idnum_clean <- ifelse(ms$idnum == "20880", ms$idnum, ms$idnum_clean)

    subset(id, idnum_clean == 36870, select = c(initials, Birthdate))
    subset(ms, idnum == "kyle36870", select = c(initials, dob))    
    ms$idnum [ms$idnum == "kyle36870"] <- 36870
    ms$idnum_clean <- ifelse(ms$idnum == "36870", ms$idnum, ms$idnum_clean)
  
    subset(id, idnum_clean == 25910, select = c(initials, Birthdate))
    subset(ms, idnum == "smith25910", select = c(initials, dob))
    ms$idnum [ms$idnum == "smith25910"] <- 25910
    ms$idnum_clean <- ifelse(ms$idnum == "25910",  ms$idnum, ms$idnum_clean)
  
    subset(ms, idnum == "grimmer0", select = c(idnum,initials,current_gr,start_gr,dob)) # MBG
    subset(id, initials == "MG") # matched on grade, and dob
    ms$idnum [ms$idnum == "grimmer0"] <- 19500
    ms$idnum_clean <- ifelse(ms$idnum == "19500",  ms$idnum, ms$idnum_clean)
  
sum(!is.na(ms$idnum_clean))
```
Now have 38 idnum cleans

#### find and correct duplicate idnum entries
```{r}
sum(duplicated(ms$idnum)) # 8

# create subset of duplicate idnums 
subset(ms, duplicated(idnum) == "TRUE", select = c(idnum, initials))

ms_dups <- subset(ms, duplicated(idnum) == "TRUE", select = c(idnum, initials, Qual_ID))

# deal with duplicates
i <- subset(ms, duplicated(idnum) == "TRUE")$idnum; i

# remove appropriate rows
    subset(ms, idnum == "18550")
        # drop 2nd entry which had no data
        ms <- subset(ms, Qual_ID != "R_2qrxvjz8mgCMqgD")
    
    subset(ms, idnum == "19600") 
        # drop 2nd entry which had no data
        ms <- subset(ms, Qual_ID != "R_3hfnvu14HLRGmcX")
    
    # 4 saw both conditions. What to do???
    subset(ms, idnum == "19400") # seems to have taken both conditions
        ms <- subset(ms, Qual_ID != "R_242WdS8eEG0ihgg") # dropping 1 row
    subset(ms, idnum == "19340") # seems to have taken both conditions but left no other data. 
        ms <- subset(ms, idnum != "19340") # dropping both data rows
    subset(ms, idnum == "43250") # seems to have taken both conditions
        ms <- subset(ms, Qual_ID != "R_1MYPt44HMZEIWza") # dropping 1 row
    subset(ms, idnum == "36510") # seems to have taken both conditions
        ms <- subset(ms, Qual_ID != "R_e2S3UBIft9mi37D") # dropping 1 row
    subset(ms, idnum == "32940") # different P each time
        # correct the idnum for 32940v1
        ms$dob <- ifelse(ms$dob == "3/30/99", "03/30/1999", ms$dob)
        subset(id, Birthdate == "03/30/1999")
        ms$idnum <- ifelse(ms$idnum == "32940" & ms$initials == "CED", "32930", ms$idnum) #update idnum
        ms$idnum_clean <- ifelse(ms$idnum == "32930", ms$idnum, ms$idnum_clean) # update idnum_clean
        # correct th idnum for 32940v2
        ms$dob <- ifelse(ms$dob == "5/10/01", "05/10/2001", ms$dob)
        subset(id, Birthdate == "05/10/2001") # 32940 is correct idnum for this P
        ms$idnum_clean <- ifelse(ms$idnum == "32940", ms$idnum, ms$idnum_clean) # update idnum_clean
    subset(ms, idnum == "43440") # same P two times in same condition
        ms <- subset(ms, Qual_ID != "R_7PP6mSwuTZtPzqh") # dropping 1 row without data
# No duplicates left in ms data
sum(duplicated(ms$idnum)) 


dim(ms)
sum(!is.na(ms$idnum_clean))
```
No more duplicate idnums, but 
__ What to do about Ps who saw both conditions???
Now have 39 cases and 40 clean ids

```{r add a variable of abbreviated initials}
# then apply the function
substrLR <- function(x){ 
  x = gsub("[.]", "", x) # remove periods
  x = gsub("[ ]", "", x) # remove spaces
  x = toupper(x)         # make all characters uppercase
  L = substr(x, 1, 1)    # extract first character on left
  R = substr(x, nchar(x), nchar(x)) # extract last character on right
  c(paste0(L,R))         # paste first and last letters together for output
  }

ms$init_2L <- substrLR(ms$initials); ms$init_2L
    # Change the one student who left his name in the initials section. 
    ms$init_2L [ms$initials == "Max Ralston"] <- "MR"; ms$init_2L
    ms$init_2L [ms$init_2L == "NANA"] <- NA; ms$init_2L
    sum(is.na(ms$init_2L)) # 12 NAs
```

#### Subsets of claimed and unclaimed idnums
```{r}
# create a variable to indicate if a variable value has been claimed in another dataset
ms$id_claimed <- 0
id$id_avail <- 1

# run this section to get a count of claimed idnumbers
ms$id_claimed [!is.na(ms$idnum_clean)] <- 1

# Claimed Count (cc) is the subset of claimed ids
cc <- merge(subset(ms, select = c(idnum_clean, id_claimed)),
            id, 
            by = "idnum_clean"); dim(cc)

# Unclaimed IDs (uc) are still available
id$id_key <- 1 # dummy var for ids in school key
ms$id_claimed [!is.na(ms$idnum_clean)] <- 1
  # remove claimed idnumbers from uc
  dim(id) # 448 cases in my key
  dim(ms) # 391 cases in my intervention
      # to be confirmed (tbc) idnumbers
      ms$tbc_id <- 1 # by default all ids are tbc
      ms$tbc_id [!is.na(ms$idnum_clean)] <- 0 # once a clean id is confirmed the id is no longer tbc
  sum(ms$tbc_id) # 353 cases to be confirmed
  
# I want to see only the idnums in the key that are still available
uc <- merge(subset(ms, tbc_id == 1,  # if the id is still to be confirmed
                   select = c(idnum)),   # include that id
                   id, by.x = "idnum", by.y = "idnum_clean",
                    all = T); dim(uc) # includes all in key and all Ps to be confirmed
uc <- subset(uc, id_key == 1); dim(uc); names(uc); dim(uc) # drop those who are not in the key and I get 448

    table(is.na(ms$idnum_clean))

# let's see who needs to be confirmed
uc; dim(uc)
# rename uc intials for merging comparisons
uc$init_2L <- uc$initials; uc$initials <- NULL
```

#### run the intials clean-up function and confirm the low hanging fruit
```{r}
# confirm id numbers of Ps whose number and init_2L match the key
ms_t <- subset(ms, is.na(idnum_clean), select = c(idnum, init_2L, dob))
names(ms_t)
names(uc)
id_init_match <- merge(ms_t, uc, by = c("idnum", "init_2L")); dim(id_init_match)
id_init_match$id_init_ck <- 1 # b/c everyone who made it into this df matched the key and the ms dataframe

# how many dead-on matches? 312 of 391
dim(ms_t) # 353
dim(id_init_match) # 312
dim(ms) # 391

# How many clean idnums? 
dim(subset(ms, !is.na(idnum_clean))) # 38
```
312 match dead-on for idnumbers and initials

_x_ update ms and uc datasets to reflect these 316 matches
```{r}
# Before merging, confirm that there are no duplicates in your keyed variable, not even NAs. 
    sum(duplicated(ms$idnum)) # 0 dups
    sum(duplicated(id_init_match$idnum)) # 0 dups

# merging
ms2 <- (merge(ms, 
          subset(id_init_match, select = c(idnum, id_init_ck)), # only includes those 312 who passed the match check for intials and idnumbers. 
          by = "idnum", 
          all = T)) # ensures that ms participants who are as yet unmatched are not dropped

dim(ms2) #391 is consistent with dim(ms)

# check to see if there are duplicates in your key variable
sum(duplicated(ms2$idnum))  # no dups

# update idnum_clean
ms2$idnum_clean <- with(ms2, ifelse(!is.na(ms2$id_init_ck), ms2$idnum, NA))
dim(subset(ms2, !is.na(idnum_clean))) # 312  

# test for duplicates
    ms_temp <- subset(ms2, !is.na(idnum_clean))
    sum(duplicated(ms_temp$idnum_clean)) # 0 uplicates exlcuding NAs

# overwrite ms with ms2
ms <- ms2
```

```{r}
# no non-NA dups
sum(duplicated(subset(ms, !is.na(idnum_clean)))) # 0 dups
describe(ms$idnum_clean)
```
Based on my tests, there is no reason to suspect that any idnumbers in April 2014 Mindset Intervention were inaccurate for the initials I am seeing. 

Currently have 313 idnum_cleans in my ms data and 78 without idnum_cleans. 

## try to figure out those without initials and idnumbers
```{r}
subset(ms, is.na(idnum_clean), select = c(idnum, initials, init_2L, dob, current_gr, start_gr))
```

### work on those with 5 digits in their idnum
```{r}
# list of Unclaimed IDs (uc) are still available
## remove ms$idnum_clean from id
ms$claimed <- 0
ms$claimed [!is.na(ms$idnum_clean)] <- 1
ms_claimed <- subset(ms, !is.na(idnum_clean), select = c(idnum_clean, claimed))

# I want to see only the idnums in the key that are still available
uc <- merge(ms_claimed, id, by = "idnum_clean", all = T); dim(uc) # 449 
uc$claimed
uc <- subset(uc, is.na(claimed)); dim(uc); names(uc) # this subset is the available idnumbers in the id dataframe
head(uc)

# current list of unclaimed ids
uc

# list of those who gave 5-digit idnum and have not been found yet. 
ms5 <- subset(ms, is.na(idnum_clean) & nchar(idnum) == 5, select = c(idnum, initials, init_2L, dob, current_gr, start_gr)); ms5
dim(ms5) # 39 

sum(duplicated(ms5$idnum)) # 0 dups
sum(duplicated(uc$idnum_clean)) # 0 dups

ms5$dob <- as.character(ms5$dob)
dim(subset(ms5, !is.na(dob))) # 30 of 39 have dobs
sort(ms5$dob)

ms5$dob[ms5$dob == "1/2/02"] <- "01/02/2002"  
ms5$dob[ms5$dob == "11/12/95"] <- "11/12/1995"
ms5$dob[ms5$dob == "1/26/99"] <- "01/26/1999"                           
ms5$dob[ms5$dob == "12-Jan-00"] <- "01/12/2000"                             
ms5$dob[ms5$dob == "2/11/98"] <- "02/11/1998"                
ms5$dob[ms5$dob == "2/14/00"] <- "02/14/2000"               
ms5$dob[ms5$dob == "2/19/98"] <- "02/19/1998"                              
ms5$dob[ms5$dob == "24-Apr-97"] <- "04/24/1997"                           
ms5$dob[ms5$dob == "30-Apr-89"] <- "04/30/1989"              
ms5$dob[ms5$dob == "4/20/98"] <- "04/20/1998"                
ms5$dob[ms5$dob == "5-Jun-98"] <- "06/05/1998"               
ms5$dob[ms5$dob == "6/15/98"] <- "06/15/1998"               
ms5$dob[ms5$dob == "6/5/99"] <- "06/05/1999"                 
ms5$dob[ms5$dob == "7-15 2001"] <- "07/15/2001"              
ms5$dob[ms5$dob == "8/2/01"] <- "08/02/2001"                 
ms5$dob[ms5$dob == "9-Mar-98"] <- "03/09/1998"              
ms5$dob[ms5$dob == "August 17,2001"] <- "08/17/2001"         
ms5$dob[ms5$dob == "december 25 1999"] <- "12/25/1999"       
ms5$dob[ms5$dob == "june 8,2000"] <- "06/08/2000"            
ms5$dob[ms5$dob == "November 4 1999"] <- "11/04/1999"       
ms5$dob[ms5$dob == "October 30th 2000"] <- "10/30/2000"      
ms5$dob[ms5$dob == "September 27th, 1997"] <- "09/27/1997"  
ms5$dob[ms5$dob == "12/19/00"] <-  "12/19/2000"
ms5$dob[ms5$dob == "2/26/15"]  <- "02/26/????"
ms5$dob[ms5$dob == "3-Dec-15"] <- "12/03/????" 
ms5$dob[ms5$dob == "8/17/97"] <- "08/17/1997"
ms5$dob[ms5$dob == "12/23/96"] <- "12/23/1996"

subset(ms5, select = c(dob))

# 24 matches
temp <- merge(ms5, uc, by.x = "idnum", by.y = "idnum_clean")
dim(temp)
names(temp)
subset(temp, select = c(idnum, init_2L, initials.y, dob, Birthdate))
# 
temp$grant_idnum_clean <- 1
```
22040, 36630, 43210 provided neither intials nor birthdate but matches an available idnumber. 
All the rest match on idnum and dob. 
So I will grant all of them their idnum as idnum_clean
```{r}
temp_merge <- subset(temp, select = c(idnum, grant_idnum_clean))
dim(ms)
dim(temp_merge)
ms2 <- merge(ms, temp_merge, by = "idnum", all = T)
dim(ms2) # all good on row and col counts

describe(ms2$idnum_clean)
ms2$grant_idnum_clean [is.na(ms2$grant_idnum_clean)] <- 0
table(ms2$grant_idnum_clean)
ms2$idnum_clean <- ifelse(ms2$grant_idnum_clean == 1, ms2$idnum, ms2$idnum_clean)
describe(ms2$idnum_clean)

ms2$grant_idnum_clean <- NULL
names(ms2)
ms <- ms2
```
Now have 337 of Ps ID'd and 54 NAs

#### Look at those with less than 5 idnums provided
```{r}
# list of Unclaimed IDs (uc) are still available
## remove ms$idnum_clean from id
ms$claimed <- 0
ms$claimed [!is.na(ms$idnum_clean)] <- 1
ms_claimed <- subset(ms, !is.na(idnum_clean), select = c(idnum_clean, claimed))
uc <- merge(ms_claimed, id, by = "idnum_clean", all = T)
uc <- subset(uc, is.na(claimed)); dim(uc); names(uc) # this subset is the available idnumbers in the id dataframe
head(uc); dim(uc) # 112 unclaimed ids
uc$Birthdate <- as.character(uc$Birthdate)

# list of those who gave less than 5-digit idnum and have not been found yet. 
subset(ms, is.na(idnum_clean) & nchar(idnum) < 5, select = c(idnum, initials, init_2L, dob, current_gr, start_gr)) # 2

subset(ms, idnum == "11")
names(id)
ms$dob <- ifelse(ms$idnum == "11", "11/31/1997", ms$dob)
"11/31/1997" %in% uc$Birthdate # not present
    # anyone in the key look like a match? 
    subset(id, initials == "MM" & Current.Grade == "11") # 19770
        # idnum 11 matches idnum 19770 on initials and Current Grade, but the birthdates don't match so I will not consider this a match.

subset(ms, idnum == "41", select = c(dob))
ms$dob <- ifelse(ms$idnum == "41", "08/11/1997", ms$dob)
"08/11/1997"  %in% uc$Birthdate # present
    # Compare to match
    subset(uc, Birthdate == "08/11/1997")
    subset(ms, idnum == "41", select = c(current_gr, init_2L))
    # update id
    ms$idnum_clean <- ifelse(ms$idnum == "41", "19950", ms$idnum_clean)
    describe(ms$idnum_clean) # 369 and 21 NAs
      

# drop excess variables
dim(ms)
names(ms)
ms <- subset(ms, select = -c(idnum_temp, idnum_ck1, idnum_verified, idnum_ck2, idcheck, claimed, id_init_ck, id_claimed, tbc_id))
names(ms)
ms <- moveme(ms, c("idnum_clean"), "before", "idnum")
ms <- moveme(ms, c("init_2L"), "after", "initials")
```

#### Who is left to confirm? 
```{r}
describe(ms$idnum_clean) # 369      21 NAs

# to be identified
tbi <- subset(ms, is.na(idnum_clean)); dim(tbi) # 21 missing idnum_clean

# merge based on dob if there are no duplicates
sum(duplicated(tbi$dob)) # 5
subset(tbi, select = c(idnum, dob, init_2L))
sort(tbi$dob)       
ms$dob <- ifelse(ms$dob == "12-Jan-00", "01/12/2000", ms$dob)    
ms$dob <- ifelse(ms$dob == "12/14/01", "12/14/2001", ms$dob)       
ms$dob <- ifelse(ms$dob == "15-Apr-98", "04/15/1998", ms$dob)      
ms$dob <- ifelse(ms$dob == "18-May-99", "05/18/1999", ms$dob)     
ms$dob <- ifelse(ms$dob == "2/19/98", "02/19/1998", ms$dob)       
ms$dob <- ifelse(ms$dob == "2001", "??/??/2001", ms$dob)           
ms$dob <- ifelse(ms$dob == "24-Apr-97", "04/24/1997", ms$dob)       
ms$dob <- ifelse(ms$dob == "28-Nov-00", "11/28/2000", ms$dob)     
ms$dob <- ifelse(ms$dob == "5-Jun-98", "06/05/1998", ms$dob)       
ms$dob <- ifelse(ms$dob == "6/4/96", "06/04/1996", ms$dob)         
ms$dob <- ifelse(ms$dob == "8/2/01", "08/02/2001", ms$dob)        
ms$dob <- ifelse(ms$dob == "9-Mar-98", "03/09/1998", ms$dob)        
ms$dob <- ifelse(ms$dob == "August 17,2001", "08/17/2001", ms$dob) 
ms$dob <- ifelse(ms$dob == "November 4 1999", "11/04/1999", ms$dob)

# recreate tbi with clean dobs
tbi <- subset(ms, is.na(idnum_clean)); dim(tbi) # 21 missing idnum_clean
dim(ms)
names(id)

id$id_avail <- NULL
id$id_key <- NULL
id$init_key <- id$initials; id$initials <- NULL
id$claimed <- id$idnum_clean %in% ms$idnum_clean
uc_ids <- subset(id, claimed == "FALSE")
dim(uc_ids) # 79 unclaimed
subset(tbi, select = c(idnum, dob, init_2L))
uc_ids

sum(duplicated(uc_ids$Birthdate)) # 0 dups
sum(duplicated(tbi$dob)) # 5 dups
    # Who are they
    subset(tbi, duplicated(dob) == "TRUE")
          # all unique IDs, none with survey data
    # none of the duplicate birthdates match the key so go ahead with merge
    temp <- merge(tbi, uc_ids, by.x = "dob", by.y = "Birthdate"); dim(temp)
    temp$idnum
    names(temp)
    subset(temp, select = c("dob","idnum_clean.y", "idnum", "current_gr", "Current.Grade", "init_2L","init_key"))
    # these 6 ids are sure matches based on dob, initials and grade
    subset(temp, select = c(idnum, idnum_clean.y))
    # upates idnum_clean in ms
    describe(ms$idnum_clean)
    ms$idnum_clean <- ifelse(ms$idnum == "28090", "38090", ms$idnum_clean)
    ms$idnum_clean <- ifelse(ms$idnum == "78830", "38870", ms$idnum_clean)
    ms$idnum_clean <- ifelse(ms$idnum == "366810", "36680", ms$idnum_clean)
    ms$idnum_clean <- ifelse(ms$idnum == "27000", "22060", ms$idnum_clean)
    ms$idnum_clean <- ifelse(ms$idnum == "25070", "21020", ms$idnum_clean)
    ms$idnum_clean <- ifelse(ms$idnum == "32600", "36470", ms$idnum_clean)
    > describe(ms$idnum_clean)
>      n missing  unique 
>    375      15     375 


#Who are my 15 holdouts? 
tbi <- subset(ms, is.na(idnum_clean), select = c(idnum, initials, current_gr, dob)); tbi

id$claimed <- id$idnum_clean %in% ms$idnum_clean
uc_ids <- subset(id, claimed == "FALSE", select = -c(Gender,Ethnicity,claimed))
dim(uc_ids) # 73 unclaimed

# are any tbi idnums among the uc_ids?
table(tbi$idnum %in% uc_ids$idnum_clean) # all false

# Look at subset of unidentified with more data
    subset(tbi, !is.na(initials))
    # 42680 LC  10 02/19/1998 is very close to 42860 02/18/1998 10  LC
    ms$idnum_clean <- ifelse(ms$idnum == "42680", "42860", ms$idnum_clean)

    # searching for 44870 JS 7 12/14/2001
    uc_ids; subset(ms, is.na(idnum_clean) & !is.na(initials), select = c(idnum, initials, current_gr, dob))
      subset(uc_ids, init_key == "JS") # nothing
      sort(uc_ids$Birthdate); subset(uc_ids, Birthdate == "12/14/1998") # nope
      # giving up on matching 44870

    # 44980      BBB          9     01/12/2000
    uc_ids; subset(ms, is.na(idnum_clean) & !is.na(initials), select = c(idnum, initials, current_gr, dob))
    sort(uc_ids$Birthdate)
    sort(uc_ids$init_key)
    # giving up on matching 44980
    
    # I see nothing on which to match those below either
        # 45040      AHW          9 11/04/1999
        # 45080      HJH          7 08/17/2001
        # 45150       wg          7 08/02/2001
        # 45180      mod          7 ??/??/2001
        # 45230      MPE         11 04/24/1997

describe(ms$idnum_clean)
```
Finished with ms data cleaning
> describe(ms$idnum_clean)
>      n missing  unique 
>    376      14     376 
__ YOU WILL JUST NEED TO MAKE A DECISION ABOUT HOW TO TREAT IDNUM_CLEANS WITH NA WHEN YOU ANALYZE DATA

## __ Make sure you didn't drop Ps who saw intervention, but didn't complete the survey, they may still show treatment effects!!!
How should you handle them? 

## MERGE MS DATA WITH EXISTING DATAFRAMES
```{r review progress on merged file}
# m is the current master merged file. 
dim(m) 

# variable order and naming clean before merger (time-stamping)
names(m)
    m <- moveme(m, c("Pattern"), "before", "Pattern1")
    
    m$startOct14 <- m$start_time; m$start_time <- NULL                                    
    m$stopOct14 <- m$stop_time; m$stop_time <- NULL
    m <- moveme(m, c("startOct14", "stopOct14"), "after", "initials")

# idnum_clean will be your key variable for the merger
describe(m$idnum_clean); class(m$idnum_clean)
m$idnum_clean <- as.character(m$idnum_clean)
sum(!is.na(m$Qual_ID)) # 417 took the survey is consistent with earlier figures. 
sum(duplicated(m$Qual_ID)) # these 48 duplicated Qual_IDs are the NAs who I have data for, but did not take survey

# ms data review
describe(ms$idnum_clean); class(ms$idnum_clean) # 376 clean idnums and 14 NAs

sum(is.na(ms$Qual_ID)) # the entire ms dataframe are survey takers 

# update ms$idnum_clean for merger
names(ms)
t <- subset(ms, is.na(idnum_clean), select = c(idnum, initials, dob)); t
    # all offered an idnum, it just didn't match my key. 
    table(t$idnum %in% ms$idnum_clean)
    # so let those be their idnum_cleans
    ms2 <- ms
    ms2$idnum_clean <- ifelse(is.na(ms2$idnum_clean), ms2$idnum, ms2$idnum_clean)
describe(ms2$idnum_clean)
    table(ms2$idnum_clean %in% ms$idnum_clean) # looks right
    sum(duplicated(ms2$idnum_clean)) # no dups
    ms <- ms2

# begin the merge
    # create temp ids for check
    as.numeric(m$idnum_clean) -> m$idnum_ck1
    as.numeric(ms$idnum_clean) -> ms$idnum_ck2
    # run preliminary checks    
    dim(m)
    dim(ms)
    sum(duplicated(m$idnum_clean)) # 0
    sum(duplicated(ms$idnum_clean)) # 0 

    # tag the ms data 
    names(ms)[2:length(names(ms))] <- paste0(names(ms)[2:length(names(ms))], "_s15")
    names(ms)

    # MERGE
    m2 <- merge(m, ms, by = "idnum_clean", all = T) 
    dim(m2)
    # check the merge
    table(ms$idnum_clean %in% m$idnum_clean)
    table(m$idnum_clean %in% ms$idnum_clean)
        # 376 of my mindset Ps are in the m dataframe
    sum(duplicated(m2$idnum_clean)) # 0 
    table(m2$idnum_ck1 - m2$idnum_ck2) # 376 0s
    summary(m2$idnum_clean) # 480 IDs total
     

# clean up the merge
names(m2)
# drop excess vars
m2 <- subset(m2, select = -c(idnum_ck2_s15, idnum_ck1, wrong.id_s15))

dim(m2)

# NOTE: on 02.10.16 it seems that the above df m2 was never integrated in the following dfs. The name m2 was used again below in a different merger.

# so on 02.10.16 I'm overwriting m2 as m so my code below remains intact
m <- m2
```
Have 480 367
  How many viable cases? 
  
```{r relabel levels of mindset manip from Spring 2015}
names(m)
# clean up mindset condition labels
levels(m$cond_ms_s15)
m$cond_ms_s15 <- as.character(m$cond_ms_s15)
m$cond_ms_s15 [m$cond_ms_s15 == "Growth (~NP15S)"] <- "Growth"
m$cond_ms_s15 <- as.factor(m$cond_ms_s15)
```

# IV. 2014-2015 Performance and Demographic Data (provided in June 2015)
```{r}
setwd("~/Box Sync/LD/10_LD_data_merging/06_2014-15_LDstudentsData")
```

## clean and cast wide the 2014-15 grades data
```{r clean and cast wide the grades data}
gr <- read.csv("Student_Academic_Final.csv", na.strings = "", head = T)
  names(gr)
  head(gr)

  # rename variables
  gr$gr_yr <- gr$Grade_In_School; gr$Grade_In_School <- NULL
  gr$score <- gr$Grade; gr$Grade <- NULL
  gr$idnum_clean <- as.character(gr$Student_ID_Transform); g$Student_ID_Transform <- NULL

  # 20 course performance grades have no course name associated with them!!!
  sum(is.na(gr$Course)); class(gr$Course)

# Prepare grades data for merging
  # paste course identyifing data into one variable course_yr_term
  gr$Course2 <- gsub("[ ]", "_", gr$Course)
  gr$Course2 <- gsub("[-]", "", gr$Course2)
  gr$Course2 <- gsub("[__]", "_", gr$Course2)
  gr$course_yr_trm <- paste(gr$Course2, "1415t", gr$Semester_Term, sep="")
  gr$course_yr_trm <- tolower(gr$course_yr_trm)

# Convert data from long to wide format using cast()
  # create subset that drops nameless courses from data!!!
  cp <- subset(gr, !is.na(Course))
  dim(cp)
  sum(is.na(cp$idnum_clean)) # 0 NAs

  # view key variables
  describe(cp$idnum_clean); length(unique(cp$idnum_clean)) # 282 unique IDs
  summary(cp$course_yr_trm)
  describe(cp$score)
  
  # create a subset of the variables you will transform to a wide dataset
  cp2 <- subset(cp, select = c(idnum_clean, course_yr_trm, score))
  dim(cp2)
  head(cp2)

# load packages
library(reshape2)
w <- dcast(cp2, idnum_clean ~ course_yr_trm,
           value.var = "score",
           fun.aggregate = mean, 
           na.rm = TRUE)
      # what is the effect of na.rm=T here???
      # what is the effect of mean in fun.aggregate???

# confirm expected number of variables and cases after casting
  describe(cp$idnum_clean); dim(w) 
  # 282 unique idnumbers is consistent across both functions 
  describe(cp$course_yr_trm)
  # 236 + 1 (for IDnum col)  matches describe(cp$course_yr_trm) values for unique course_yr_trm values

# prep for merging with main dataframe
dim(m)

# tag the cp data 
    names(cp)[2:length(names(cp))] <- paste0(names(cp)[2:length(names(cp))], "_1415")
    names(cp)

# test vars for merger
w$testw <- as.numeric(w$idnum_clean)
m$testm <- as.numeric(m$idnum_clean)

m2 <- merge(m, w, by = "idnum_clean", all = TRUE); dim(m2) # 512

m2$test <- m2$testw - m2$testm; table(m2$test) # 236 0s
m2$test <- NULL; m2$testw <- NULL; m2$testm <- NULL

# merger looks good, so overwrite
m <- m2 # 
dim(m)
```

## clean progress testing data
```{r clean progress testing data}
prog <- read.csv("Student_ProgressTesting_Final.csv", na.strings = "", head = T)
    dim(prog); names(prog) # 450  24
    # clean up names of vars
    prog$Pattern1415 <- prog$Pattern_new; prog$Pattern_new <- NULL
    prog$GORT_date1415 <- prog$GORT_TestDate_new; prog$GORT_TestDate_new <- NULL 
    prog$idnum_clean <- as.numeric(prog$StudentID); prog$StudentID <- NULL
      prog <- moveme(prog, c("idnum_clean"), "before", "DSM_1")
    sum(is.na(prog$idnum_clean)) # 0 NAs
    sum(duplicated(prog$idnum_clean)) # 2 duplicates
        # Correct duplicates
        subset(prog, duplicated(idnum_clean) == "TRUE", select = idnum_clean)
            # 17590 drop the second entry after averaging his GORT scores
            t <- subset(prog, idnum_clean == "17590"); t
              # create new row averaging GORT scores
              prog$GORT_RC <- ifelse(prog$idnum_clean == "17590", mean(t$GORT_RC), prog$GORT_RC)
              prog$GORT_RR <- ifelse(prog$idnum_clean == "17590", mean(t$GORT_RR), prog$GORT_RR)  
              # drop second entry for 17590
              prog$idnum_clean <- ifelse(prog$idnum_clean == "17590" & prog$DSM_1 == 315.00, "17590drop", prog$idnum_clean)
              prog <- subset(prog, idnum_clean != "17590drop")
              dim(prog)
            # 33420
              t <- subset(prog, idnum_clean == "33420"); t
              # 33420 has two sets of scores, keep the more recent WISC, and average GORTs
              prog$GORT_RA <- ifelse(prog$idnum_clean == "33420", mean(t$GORT_RA), prog$GORT_RA)
              prog$GORT_RC <- ifelse(prog$idnum_clean == "33420", mean(t$GORT_RC), prog$GORT_RC)
              prog$GORT_RR <- ifelse(prog$idnum_clean == "33420", mean(t$GORT_RR), prog$GORT_RR) 
              # drop second entry for 33420
              prog$idnum_clean <- ifelse(prog$idnum_clean == "33420" & prog$PerceptualReasoning == 94, "33420drop", prog$idnum_clean)
              prog <- subset(prog, idnum_clean != "33420drop")
              dim(prog) # 448 24
```

```{r Sort by Pattern (~text-to-column)}

library(splitstackshape)
dim(prog)
t <- subset(prog, select = c(idnum_clean, Pattern1415)); dim(t) # 466   2
t2 <- cSplit(t, 'Pattern1415', sep=", ", type.convert=FALSE)
names(t2) # no one had more than 6 Patterns

i <- 1
t2$Pattern1_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern1_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern1_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern1_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern1_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern1_1415 [t2$Pattern1415_6 == i] <- i

i <- 2
t2$Pattern2_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern2_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern2_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern2_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern2_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern2_1415 [t2$Pattern1415_6 == i] <- i

i <- 3
t2$Pattern3_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern3_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern3_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern3_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern3_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern3_1415 [t2$Pattern1415_6 == i] <- i

i <- 4
t2$Pattern4_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern4_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern4_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern4_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern4_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern4_1415 [t2$Pattern1415_6 == i] <- i

i <- 5
t2$Pattern5_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern5_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern5_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern5_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern5_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern5_1415 [t2$Pattern1415_6 == i] <- i

i <- 6
t2$Pattern6_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern6_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern6_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern6_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern6_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern6_1415 [t2$Pattern1415_6 == i] <- i

i <- 7
t2$Pattern7_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern7_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern7_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern7_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern7_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern7_1415 [t2$Pattern1415_6 == i] <- i

i <- 8
t2$Pattern8_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern8_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern8_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern8_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern8_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern8_1415 [t2$Pattern1415_6 == i] <- i

i <- 9
t2$Pattern9_1415 [t2$Pattern1415_1 == i] <- i
t2$Pattern9_1415 [t2$Pattern1415_2 == i] <- i
t2$Pattern9_1415 [t2$Pattern1415_3 == i] <- i
t2$Pattern9_1415 [t2$Pattern1415_4 == i] <- i
t2$Pattern9_1415 [t2$Pattern1415_5 == i] <- i
t2$Pattern9_1415 [t2$Pattern1415_6 == i] <- i

head(t2)
describe(t2$Pattern1_1415)
describe(t2$Pattern2_1415)
describe(t2$Pattern3_1415)
describe(t2$Pattern4_1415)
describe(t2$Pattern5_1415)
describe(t2$Pattern6_1415)
describe(t2$Pattern7_1415)
describe(t2$Pattern8_1415)
describe(t2$Pattern9_1415)

# Remove variables you don't want merged back in with the main dataset
t2$Pattern1415_1 <- NULL
t2$Pattern1415_2 <- NULL
t2$Pattern1415_3 <- NULL
t2$Pattern1415_4 <- NULL
t2$Pattern1415_5 <- NULL
t2$Pattern1415_6 <- NULL
names(t2)

class(t2)
t2 <- data.frame(t2)
dim(t2); colnames(t2); class(t2)
```

```{r merge with prog, the survey data with complete idnumbers}
prog$testidm <- as.numeric(prog$idnum_clean)
t2$testidt <- as.numeric(t2$idnum_clean)
prog2 <- merge(prog, t2, by = "idnum_clean", all = TRUE); dim(m) # 
  # test merger
  prog2$test <- prog2$testidm - prog2$testidt; table(prog2$test)
  prog2$test <- NULL; prog2$testidm <- NULL; prog2$testidt <- NULL
  # check the merger
  length(unique(prog2$idnum_clean)) # 
  sum(duplicated(prog2$idnum_clean)) # 0
  sum(is.na(prog2$idnum_clean)) # 0

dim(prog2)  # 448  33
names(prog2)

# prep for merge
    m$idnum_ck1 <- as.numeric(m$idnum_clean)
    dim(m); sum(duplicated(m$idnum_clean))
    
    prog2$idnum_ck2 <- as.numeric(prog2$idnum_clean)
    dim(prog2); sum(duplicated(prog2$idnum_clean))

    # tag the prog data 
    names(prog2)[2:length(names(prog2))] <- paste0(names(prog2)[2:length(names(prog2))], "_1415")
    names(prog2)

# merge
    m2 <- merge(m, prog2, by = "idnum_clean", all = T)
    dim(m2) # column counts are good, 13 new Ps entered my data
    table(m2$idnum_ck1 - m2$idnum_ck2) # 435 0s

m <- m2
dim(m) # 525 549
names(m)
```

```{r create composite variables from progress testing}
# WISC IQ
m$WISC_1415 <- apply(X = m[,c("WorkingMemory_1415", "VerbalComp_1415", "PerceptualReasoning_1415", "ProcessingSpeed_1415")], MARGIN =1, FUN = mean, na.rm = T)
m <- moveme(m, c("WISC_1415"), "before", "WorkingMemory_1415")
  label(m$WISC_1415) = "mean WISC (IQ) score from 2014-15 data" 
  describe(m$WISC_1415)
  hist(m$WISC_1415)

# GORT
m$gort_1415 <- apply(X = m[,c("GORT_RA_1415",  "GORT_RC_1415",  "GORT_RR_1415")], MARGIN =1, FUN = mean, na.rm = T)
m <- moveme(m, c("gort_1415"), "before", "GORT_RA_1415")
  label(m$gort_1415) = "mean GORT score from 2014-15 data" 
  describe(m$gort_1415)
  hist(m$gort_1415)
```

## clean demo data
```{r clean demo data}
dem <- read.csv("Student_Demographic_Final.csv", na.strings = "", head = T)
    dim(dem); names(dem)   # 450 5
    dem$idnum_clean <- as.character(dem$Student_ID); dem$StudentID <- NULL
    sum(is.na(dem$idnum_clean)) # 0 NAs
    sum(duplicated(dem$idnum_clean)) # 0 duplicates

# begin merging
  # temp IDs for testing the merge
  m$test_id_m <- as.numeric(m$idnum_clean)
  dem$test_id_dem <- as.numeric(dem$idnum_clean)

    # tag the dem data 
    names(dem)[2:length(names(dem))] <- paste0(names(dem)[2:length(names(dem))], "_1415")
    names(dem)
  # merge
  m2 <- merge(m, dem, by = "idnum_clean", all = TRUE)
  dim(m2) # 525
      # test merge
      m2$test <- m2$test_id_m - m2$test_id_dem; table(m2$test) # all zeroes confirms clean merge
      # drop testing vars
      m2$test <- NULL; m2$test_id_m <- NULL;  m2$test_id_dem <- NULL 
  
  # confirm no NAs or duplicates for idnum_clean
  sum(is.na(m2$idnum_clean))
  sum(duplicated(m2$idnum_clean)) # 0

# overwrite m2 with m
m <- m2
dim(m) # 525 554
names(m)
```

# V. October 2015 survey data
```{r load Oct 2015 data}
setwd("~/Box Sync/LD/10_LD_data_merging/07_Oct2015_LDsurvey")
f2 <- read.csv('LD_Mindset_Oct2015_master.csv', head=T)
dim(f2)
names(f2) 

# drop the timing variables
f2 <- subset(f2, select = -c(Intro1, Q288_1,   Q288_2, 	Q288_3, 	Q288_4, 	Q219_1, 	Q219_2, 	Q219_3, 	Q219_4, 	Q221_1, 	Q221_2, 	Q221_3, 	Q221_4, 	Q222_1, 	Q222_2, 	Q222_3, 	Q222_4, 	Q223_1, 	Q223_2, 	Q223_3, 	Q223_4, 	Q224_1, 	Q224_2, 	Q224_3, 	Q224_4, 	Q226_1, 	Q226_2, 	Q226_3, 	Q226_4, 	Q225_1, 	Q225_2, 	Q225_3, 	Q225_4, 	Q174_1, 	Q174_2, 	Q174_3, 	Q174_4, 	Q173_1, 	Q173_2, 	Q173_3, 	Q173_4, 	Q175_1, 	Q175_2, 	Q175_3, 	Q175_4, 	Q271_1, 	Q271_2, 	Q271_3, 	Q227_1, 	Q227_2, 	Q227_3, 	Q227_4, 	Q265_1, 	Q265_2, 	Q265_3, 	Q265_4, 	Q266_1, 	Q266_2, 	Q266_3, 	Q266_4, 	Q267_1, 	Q267_2, 	Q267_3, 	Q267_4, 	Q268_1, 	Q268_2, 	Q268_3, 	Q268_4, 	Q269_1, 	Q269_2, 	Q269_3, 	Q269_4, 	Q270_1, 	Q270_2, 	Q270_3, 	Q270_4, 	Q271_1, 	Q271_2, 	Q271_3, 	Q271_4, 	Q272_1, 	Q272_2, 	Q272_3, 	Q272_4, 	Q273_1, 	Q273_2, 	Q273_3, 	Q273_4, 	Q274_1, 	Q274_2, 	Q274_3, 	Q274_4, 	Q291_1, 	Q291_2, 	Q291_3, 	Q291_4, 	Q275_1, 	Q275_2, 	Q275_3, 	Q275_4, 	Q276_1, 	Q276_2, 	Q276_3, 	Q276_4, 	Q278_1, 	Q278_2, 	Q278_3, 	Q278_4, 	Q280_1, 	Q280_2, 	Q280_3, 	Q280_4, 	Q292_1, 	Q292_2, 	Q292_3, 	Q292_4, 	Q281_1, 	Q281_2, 	Q281_3, 	Q281_4, 	Q282_1, 	Q282_2, 	Q282_3, 	Q282_4, 	Q283_1, 	Q283_2, 	Q283_3, 	Q283_4, 	Q284_1, 	Q284_2, 	Q284_3, 	Q284_4, 	Q293_1, 	Q293_2, 	Q293_3, 	Q293_4, 	Q286_1, 	Q286_2, 	Q286_3, 	Q286_4, 	Q294_1, 	Q294_2, 	Q294_3, 	Q294_4, 	Q295_1, 	Q295_2, 	Q295_3, 	Q295_4, LocationLatitude, LocationLongitude, LocationAccuracy))
f2$IDS <- NULL # this is the vector of possible idnumbers
dim(f2)

# f2$idnum_clean <- NULL (not sure why this variable was dropped) 02.10.16
f2$idnum_orig <- as.character(f2$idnum_orig)

# do idnum_clean and idnum_orig match in this dataset? 
dim(subset(f2, idnum_clean != idnum_orig, select = idnum_orig)) # 46 don't match
sort(subset(f2, idnum_clean != idnum_orig, select = idnum_orig)$idnum_orig)

# remove punctuation and letters from data
f2$idnum <- gsub('[[:alpha:]]|[[:punct:]]','',f2$idnum_orig) #

# add the 0 ot the end of 4 digit idnumbers so they'll match your key. 
f2$idnum <- ifelse(nchar(f2$idnum) == 4, 
                            paste0(f2$idnum, "0"), 
                            f2$idnum)

# review for odd length idnums
sort(f2$idnum)
  # !!! idnums to be sorted out!!!
  subset(f2, idnum == 0, select = c(initials, current_gr, start_gr, dob))
  subset(f2, idnum == 415, select = c(initials, current_gr, start_gr, dob))

  describe(f2$idnum)
  length(unique(f2$idnum)) # 328 cases, 324 unique, 0 NAs
  sum(duplicated(f2$idnum)) # 4 dups
  
# Resolve these duplicates
  subset(f2, duplicated(idnum) == T, select = idnum)
  # these cases are being dropped b/c they are clearly logout errors
  subset(f2, idnum == "42500")
    f2$idnum [f2$idnum == "42500" & f2$initials == ""] <- "drop"
  subset(f2, idnum == "45910")
    f2$idnum [f2$idnum == "45910" & f2$initials == ""] <- "drop"
  subset(f2, idnum == "19660")  
    f2$idnum [f2$idnum == "19660" & f2$initials == "CMK" ] <- "19610"
  subset(f2, idnum == "46580")  
    f2$idnum [f2$idnum == "46580" & f2$initials == ""] <- "drop"

# idnum_clean = 0 needs to be identified!!!
subset(f2, idnum_clean == 0)

# this person entered idnum orig of 415 and I can't identify them
f2$idnum [f2$idnum == "41500" & f2$initials == "ahw" ] <- "drop"   
    # no initials for AHW, AW initials already claimed by AJW

# 4 rows need to be dropped, not just NA'd!!!
names(f2)

# view the three idnum variables
subset(f2, select = c(idnum_clean, idnum_orig, idnum)))

# drop those you've deemed droppable
f <- subset(f2, idnum != "drop")

f$Oct2015taker <- 1
f$idnum_clean

# I'm not sure who idnum_clean = 0 is so I'm dropping them for now. 
f <- subset(f, idnum_clean != "0")

# rename and drop dead vars
names(f)
f$Qual_ID_F15 <- f$V1; f$V1 <- NULL
f$start <- f$V8
f$stopped <- f$V9
f <- subset(f, select = -c(V2, V3,  V4,	V5,	V6,	V7, V8, V9, V10, wrong.id))

# label all data in f with year
names(f)
names(f)[2:length(names(f))] <- paste0(names(f)[2:length(names(f))], "_f15")
```

## merge f and m
```{r}
m2 <- merge(m,f, by = "idnum_clean", all = T)
dim(m2)
names(m2)

m2 <- subset(m2, select = -c(Q271_1.1_f15:Q271_4.1_f15))

# If m2 looks good, overwrite it as m
names(m2)
m <- m2
```
Every variable appears to be time stamped now. 

## order your variables wisely
```{r}
# generate list of column or variable names without punctuation
m[0,] 

# Drop the excess variables
m <- subset(m, select = -c(idnum_orig_f14:ins_clk_f14, GRins_f14, debrief_f14, dwnld_f14))

m <- subset(m, select = -c(tempIDdup_check_f14))

m <- subset(m, select = -c(test_id_dem_1314, test_id_prog_1314, idnum_s15, Qual_ID_s15, initials_s15, init_2L_s15, vid_s15, vid_play_s15, image_s15))

m <- subset(m, select = -c(image_TEXT_s15, RC_ins_s15, idnum_orig_s15, idnum_ck2_1415, idnum_orig_f15, initials_f15, Qual_ID_F15_f15, Q216_f15, Q217_f15, Q218_f15, idnum_f15))    

dim(m)

# Create new variables for organizing
m$BRK_______________DEMOGnPERFORMANCE_201314_DATA_______________BRK <- NA
m$BRK_______________FALL2014_DATA_______________BRK <- NA
m$BRK_______________SPRING2015MS_DATA_______________BRK <- NA
m$BRK_______________DEMOGnPERFORMANCE_201415_DATA_______________BRK <- NA
m$BRK_______________FALL2015_DATA_______________BRK <- NA


m <- moveme(m, c("BRK_______________DEMOGnPERFORMANCE_201314_DATA_______________BRK"), "before", "Birthdate_1314") 
m <- moveme(m, c("BRK_______________FALL2014_DATA_______________BRK"), "before", "B_E1_f14") 
m <- moveme(m, c("BRK_______________SPRING2015MS_DATA_______________BRK"), "before", "cond_ms_s15")
m <- moveme(m, c("BRK_______________DEMOGnPERFORMANCE_201415_DATA_______________BRK"), "before", "6th_elective_block1415t0")
m <- moveme(m, c("BRK_______________FALL2015_DATA_______________BRK"), "before", "current_gr_f15")

m[0,] 
dim(m)

m <- subset(m, select=c(idnum_clean,             
                        BRK_______________DEMOGnPERFORMANCE_201314_DATA_______________BRK:Yearbook_1314t2, 
                        BRK_______________FALL2014_DATA_______________BRK: howfeel_oe_f14,
                        startOct14:stopOct14,
                        BRK_______________SPRING2015MS_DATA_______________BRK: anyelse_s15, 
                        BRK_______________DEMOGnPERFORMANCE_201415_DATA_______________BRK: Pattern9_1415_1415, 
                        BRK_______________FALL2015_DATA_______________BRK: stopped_f15
                        ))

m[0,] 
dim(m)
```
Dataframe m is now well organized with clear section breaks on data

## Who missed the Fall 2015 survey? 
```{r who missed Fall 2015 survey}
# compare Fall 2015 survey takers to the master list of idnumbers. 
setwd("~/Box Sync/LD/10_LD_data_merging")
# Load the key you created of school provided initials and idnums
IDinit <- read.csv('03_Oct2014_Student_Initials_ID.csv', head = T)
dim(IDinit) # 448 students and initials
# How many in my grades data? 
setwd("~/Box Sync/LD/10_LD_data_merging/04_2013-14_LDstudentsData")
dem <- read.csv("Student_Demographic_FINAL.csv", head = T); dim(dem) # 448
names(dem)
dem$idnum_clean <- dem$StudentID 
dem$Grade1415 <- dem$Current.Grade

  summary(duplicated(dem$idnum_clean)) # no duplicates in this dataframe
  dem2 <- subset(dem, Grade1415 != 12); dim(dem2) # 394 now that seniors are gone
dim(f) # so 70 students are missing off the bat due to graduation from HS

# Now perform a vlookup across the two datasets
setwd("~/Box Sync/LD/10_LD_data_merging")
# write.csv(f, 'oct2015.csv')

# merge subsets to see grades of survey nontakers for October 2015
t = merge(f[, c("idnum_clean", "current_gr", "Oct2015taker")], 
              dem2[, c("idnum_clean", "Grade1415")], 
              all = T)
dim(t)
dim(f)
dim(dem)
head(t, 10)
names(t)

# subset of Ps in demographic data (graduates already removed) who are not in my survey data
nontakers <- subset(t, is.na(Oct2015taker), select = c(idnum_clean, Grade1415))
dim(nontakers) # 106 people in my demographic data did not take the survey. 
table(nontakers$Grade1415)
# the distribution is even across grades
#write.csv(nontakers, 'nontakersOct2015.csv')

# confirm how many are in the Fall 2015 dataset that weren't in old data. 
```

### Verify idnum_clean for cases whose intials and idnum match the key for Fall 2015 (f2)
```{r verify initials and idnums}
# visually review the initials
subset(f2, select = c(idnum_orig, initials))

    # visual inspection revealed long text responses from this P
    subset(f2, idnum_orig == 39470)
    # remove dob and initials for 39470
    f2$initials [f2$idnum_orig == "39470"] <- NA
    f2$dob [f2$idnum_orig == "39470"] <-  NA
    # create var for 2 letter initials
    f2$init_2L <- substrLR(f2$initials); f2$init_2L
        f2$init_2L [f2$init_2L == "NANA"] <- NA
        f2$init_2L [f2$init_2L == ""] <- NA


names(id); names(f2)
idt <- subset(id, select = c(idnum_clean, init_key))
idt$idnum <- idt$idnum_clean; idt$idnum_clean <- NULL
idt$init_2L <- idt$init_key; idt$init_key <- NULL
names(idt)

# create subset (t) of those who match on idnum and initials 
t <- merge(subset(f2, select = c(idnum, init_2L)), idt, by = c("idnum", "init_2L")); dim(t) # t now holds all 266 Ps whose cleaned idnums and initials matched those found in the key.

# create varible in f2 to indicate if case id in subset of matched initials and idnums
f2$idnum_ver <- f2$idnum %in% t$idnum

# update idnum_cleans for Ps with verified initials and idnums
f2$idnum_clean <- NA
f2$idnum_clean <- ifelse(f2$idnum_ver == "TRUE", f2$idnum, f2$idnum_clean)

# Update the key (id) to indicate if idnum has been claimed in idnum_clean
names(id)
id$claimed <- 0
id$claimed <- id$idnum_clean %in% f2$idnum_clean # indicates if id is claimed                 
```

### Birthdate conversion for October 2015 dataset (f2)
```{r dob subset}
# create a new subset of f2 Ps without an idnum_clean
        f2$dob <- as.character(f2$dob)
        f2$dob_orig <- f2$dob
  
    t <- subset(f2, is.na(idnum_clean), select = c(idnum, init_2L, initials, dob_orig, dob, current_gr))
    sort(t$dob) 
```

```{r Parse your dobs by style they are written in} 
# Parse your dobs by style they are written in:
# ##/##/##
t$dob_slash <- ifelse( grepl("/", t$dob) == "TRUE",  t$dob, NA)
t$dob_day_mon_yr <-   ifelse( grepl("[[:alpha:]]", t$dob) == "TRUE",  t$dob, NA)
t$dob_mon_day_yr <- ifelse( grepl("[[:alpha:]]", substr(t$dob, 1, 1)) == "TRUE", t$dob, NA) # tests if FIRST letter of vector element is letter
  # remove t$dob_mon_day_yr cases from t$dob_day_mon_yr
  t$dob_day_mon_yr <- ifelse(!is.na(t$dob_mon_day_yr), NA, t$dob_day_mon_yr)

# write as a function called happy_birthday() !!!
subset(t, !is.na(dob_slash), select = c(dob_slash))
subset(t, !is.na(dob_day_mon_yr), select = c(dob_day_mon_yr))
subset(t, !is.na(dob_mon_day_yr), select = c(dob_mon_day_yr))
```

```{r slash dates}
# clean them up
## dob_slash
subset(t, !is.na(dob_slash), select = c(dob_slash))
t <- cSplit(t, "dob_slash", sep = "/", type.convert = F)
  # month
  t$dob_slash_1 <- ifelse(nchar(t$dob_slash_1) == 1, paste0("0",t$dob_slash_1), t$dob_slash_1)
  subset(t, !is.na(dob_slash_1), select = c(dob_slash_1)) # visual confirmation
  t$month <- t$dob_slash_1; t$dob_slash_1 <- NULL
  # day
  subset(t, !is.na(dob_slash_2), select = c(dob_slash_2))
  t$dob_slash_2 <- ifelse(nchar(t$dob_slash_2) == 1, paste0("0",t$dob_slash_2), t$dob_slash_2)
  subset(t, !is.na(dob_slash_2), select = c(dob_slash_2)) # visual confirmation
  t$day <- t$dob_slash_2; t$dob_slash_2 <- NULL
  # yr
  subset(t, !is.na(dob_slash_3), select = c(dob_slash_3))
  t$dob_slash_3 <- ifelse(nchar(t$dob_slash_3) == 2 & t$dob_slash_3 < 15, paste0("20",t$dob_slash_3), t$dob_slash_3)
  t$dob_slash_3 <- ifelse(nchar(t$dob_slash_3) == 2 & t$dob_slash_3 > 15, paste0("19",t$dob_slash_3), t$dob_slash_3)
  subset(t, !is.na(dob_slash_3), select = c(dob_slash_3)) # visual confirmation
  t$yr <- t$dob_slash_3; t$dob_slash_3 <- NULL

  ### Paste together new bday var (dob_cln)
    t$dob_cln <- NA
    t$dob_cln <- with(t, paste(month, day, yr, sep = "/"))
    t$dob_cln <- ifelse(t$dob_cln == "NA/NA/NA", NA, t$dob_cln)
    describe(t$dob_cln) # added 26 clean birthdates to vector
    #### drop excess vars
    t <- subset(t, select = -c(month, day, yr)); names(t)
```

```{r day_mon_yr}
## dob_day_mon_yr
subset(t, !is.na(dob_day_mon_yr), select = c(dob_day_mon_yr))
### make all seperators consistent within the vector elements
  t <- cSplit(t, "dob_day_mon_yr", sep = "-", type.convert = F)
  names(t)
  # dob_day_mon_yr_1
  subset(t, !is.na(dob_day_mon_yr_1), select = c(dob_day_mon_yr_1)) # visual confirmation
  t$day <- t$dob_day_mon_yr_1; t$dob_day_mon_yr_1 <- NULL
  t$day <- with(t, ifelse(nchar(day) == 1, paste0("0", day), day)) 
  subset(t, !is.na(day), select = c(day)) # visual confirmation
  # dob_day_mon_yr_2
  subset(t, !is.na(dob_day_mon_yr_2), select = c(dob_day_mon_yr_2)) # visual confirmation
  t$month <- tolower( # make all values lowercase
    substr(t$dob_day_mon_yr_2, 1, 3)) # take only first three letters of values
    t$dob_day_mon_yr_2 <- NULL
      # create df of numbers and months
      mo_num <- c("01","02","03","04","05","06","07","08","09","10","11","12")
      month <- c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
      num_months <- data.frame(mo_num, month)
  dim(t); names(t)
  # add var to indicate num_months that you can drop
  t$tp <- 1 # as in row represents a participant (p) in df (t)
  t2 <- merge(t, num_months, by = "month", all = T); dim(t2)
    # drop the months that were merged in without a participant to match
      t2$tp # the NAs are months without participants
      t2 <- subset(t2, !is.na(tp)); t2$tp
  subset(t2, !is.na(month), select = c(month, mo_num)) # visual confirmation
  t <- t2

  # dob_day_mon_yr_3
  subset(t, !is.na(dob_day_mon_yr_3), select = c(dob_day_mon_yr_3)) # visual confirmation
  t$yr <- t$dob_day_mon_yr_3; t$dob_day_mon_yr_3 <- NULL
  t$yr <- with(t, ifelse(yr > 16, paste0("19", yr), yr)) # must run this line... 
  t$yr <- with(t, ifelse(yr < 16, paste0("20", yr), yr)) # ... before this line
  subset(t, !is.na(yr), select = c(yr)); describe(t$yr) # visual confirmation

  ### add to new bday var (dob_cln)
    describe(t$dob_cln) # have 26 dob_cln before the upcoming paste
    t$dob_cln <- ifelse(is.na(t$dob_cln), paste(t$mo_num, t$day, t$yr, sep = "/"), t$dob_cln)
    t$dob_cln <- ifelse(t$dob_cln == "NA/NA/NA", NA, t$dob_cln)
        describe(t$dob_cln) # have 44 dob_cln after the paste
        subset(t, select = c(idnum, dob_cln, dob)) # visual inspection     
    #### drop excess vars
    t <- subset(t, select = -c(mo_num, month, day, yr)); names(t)
```

```{r mon_day_yr}
subset(t, !is.na(dob_mon_day_yr), select = c(dob_mon_day_yr)) # visual confirmation

# remove excess punctuation and spaces
t$dob_mon_day_yr <- gsub(",", " ", t$dob_mon_day_yr)
t$dob_mon_day_yr <- gsub("th", "", t$dob_mon_day_yr)
t$dob_mon_day_yr <- gsub("  ", " ", t$dob_mon_day_yr)
subset(t, !is.na(dob_mon_day_yr), select = c(dob_mon_day_yr)) # visual confirmation
# text-to-col the dob
t <- cSplit(t, "dob_mon_day_yr", sep = " ", type.convert = F); names(t)

  # dob_mon_day_yr_1
  subset(t, !is.na(dob_mon_day_yr_1), select = c(dob_mon_day_yr_1)) # visual confirmation
  t$month <- tolower( # make all values lowercase
                    substr(t$dob_mon_day_yr_1, 1, 3)) # take first three letters
                    t$dob_mon_day_yr_1 <- NULL
  # create df of numbers and months
  mo_num <- c("01","02","03","04","05","06","07","08","09","10","11","12")
  month <- c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
  num_months <- data.frame(mo_num, month)
  dim(t); names(t)
  t$tp <- 1 # as in row represents a participant (p) in df (t)
  t2 <- merge(t, num_months, by = "month", all = T); dim(t2)
  # drop the months that were merged in without a participant to match
  t2$tp # the NAs are months without participants
  t2 <- subset(t2, !is.na(tp)); t2$tp; dim(t2)
  subset(t2, !is.na(month), select = c(month, mo_num)) # visual confirmation
  t <- t2

  # dob_mon_day_yr_2
  subset(t, !is.na(dob_mon_day_yr_2), select = c(dob_mon_day_yr_2)) # visual confirmation
  t$day <- t$dob_mon_day_yr_2; t$dob_mon_day_yr_2 <- NULL
  t$day <- with(t, ifelse(nchar(day) == 1, paste0("0", day), day)) 
  subset(t, !is.na(day), select = c(day)) # visual confirmation
  
  # dob_mon_day_yr_3
  subset(t, !is.na(dob_mon_day_yr_3), select = c(dob_mon_day_yr_3)) # visual confirmation
  t$yr <- t$dob_mon_day_yr_3; t$dob_mon_day_yr_3 <- NULL
  subset(t, !is.na(yr), select = c(yr)) # visual confirmation

  ### add to new bday var (dob_cln)
    describe(t$dob_cln) # have 44 dob_cln before the upcoming paste
    t$dob_cln <- ifelse(is.na(t$dob_cln), paste(t$mo_num, t$day, t$yr, sep = "/"), t$dob_cln)
    t$dob_cln <- ifelse(t$dob_cln == "NA/NA/NA", NA, t$dob_cln)
        t$dob_cln # visual inspection
        describe(t$dob_cln) # have 52 dob_cln after the paste
    #### drop excess vars
    t <- subset(t, select = -c(month, day, yr, tp, mo_num)); names(t)
    dim(t)
```
of the 60 Ps without an idnum_clean 52 now have clean dob
See who you can match among those 52 Ps to the key
```{r See who you can match among those 52 Ps to the key} 
names(id)
idt <- subset(id, claimed == "FALSE")

sum(duplicated(idt$Birthdate)) # 12 dups

# create var indicating which unconfirmed Ps in your data among the unclaimed in the key
t$id_verif <- t$dob_cln %in% idt$Birthdate
# create var indicating which unclaimed in the key are among the unconfirmed Ps in your data 
idt$id_verif <- idt$Birthdate  %in%  t$dob_cln

    # view those subsets of matches
    t_t <- subset(t, id_verif == T, select = c(idnum, init_2L, current_gr, dob_cln))
    sum(duplicated(t_t$dob_clean))

    idt_t <- subset(idt, id_verif == T, select = c(idnum_clean, init_key, Current.Grade, Birthdate))

    # check for dups in key var b4 merging
    sum(duplicated(idt_t$Birthdate)) # 2 dups
    subset(idt_t, duplicated(Birthdate) == T) # 01/03/2001
    subset(idt_t, Birthdate == "01/03/2001")
      # who do they match in t_t
      subset(t_t, dob_cln == "01/03/2001")
      subset(idt, idnum_clean == "38430")
      # so 38430 is confirmed by idnum and birthdate
      # update f2$idnum_clean
      describe(f2$idnum_clean)
      f2$idnum_clean <- with(f2, ifelse(idnum == "38430", idnum, idnum_clean))
      describe(f2$idnum_clean)
      # update idt and id$claimed
      id$claimed <- id$idnum_clean %in% f2$idnum_clean # indicates if id is claimed                 
      idt <- subset(id, claimed == F); dim(idt) # 181 unclaimed
      # update t_t
      names(t_t); dim(t_t)
      t_t <- subset(t_t, idnum != "38430")

idt$dob_cln <- idt$Birthdate
idt_t <- subset(idt, select = c(dob_cln, idnum_clean, Current.Grade, init_key))
# update Current.Grade for this year by adding 1
idt_t$Current.Grade <- idt_t$Current.Grade + 1

sum(duplicated(t_t$dob_cln)) # 0 dups
sum(duplicated(idt_t$dob_cln)) # 11 dups
    # do any of these dups create conflicts? 
    sum(duplicated(idt_t$dob_cln)) # 11 dup dobs among unclaimed cases in key
    sum(subset(idt_t, duplicated(dob_cln) == T)$dob_cln %in% t_t$dob_cln) # but none are in t_t
    # proceed with merger
    temp <- merge(t_t, idt_t, by = "dob_cln", all.x = T)
    dim(t_t); dim(temp) # all t_t were maintained

temp$init_ck <- temp$init_2L == temp$init_key
temp$id_ck <- temp$idnum == temp$idnum_clean
temp$gr_ck <- temp$current_gr == temp$Current.Grade
temp

# verified idnum_clns
## these people match on dob and idnum; that's a sure match
temp$id_verif <- with(temp, ifelse(id_ck == T, idnum, NA))
subset(temp, !is.na(id_verif))
      describe(f2$idnum_clean) # n = 269  59 NAs but there are 2 dups
          #subset(f2, duplicated(idnum_clean) == T)$idnum_clean  # 42500 & 19660
          # subset(f2, idnum_clean == "42500")
          #f2 <- subset(f2, V1 != "R_1DA75zGJrRA3sWo")# dropping empty row for 42500
      f2$idnum_verified <- f2$idnum %in% temp$id_verif
      describe(f2$idnum_verified) # 17 matches
      f2$idnum_clean <- with(f2, ifelse(is.na(idnum_clean) & idnum_verified == T, idnum, idnum_clean))
      describe(f2$idnum_clean) # n = 286, an increase of 17

## these people match on dob, initials, & grade; that's a sure match
temp$init_verif <- with(temp, ifelse(init_ck == T, idnum_clean, NA))
subset(temp, !is.na(init_verif))
      describe(f2$idnum_clean) # n = 269      59 NAs
      f2$idnum_verified <- f2$idnum %in% temp$id_verif
      describe(f2$idnum_verified) # 17 matches
      f2$idnum_clean <- with(f2, ifelse(is.na(idnum_clean) & idnum_verified == T, idnum, idnum_clean))
      describe(f2$idnum_clean) # n = 286, an increase of 17






## who's left of the dob matches?
temp$judg_verif <- NA
subset(temp, is.na(id_verif) & is.na(init_verif))
# 38430 # only matches on dob; unidentifiable!!! test against newer grades data
temp$judg_verif <- with(temp, ifelse(idnum == "24900", "42490", temp$judg_verif)) # id only off by 1
# 45210 # only matches on dob; unidentifiable!!! test against newer grades data
# 39430 # only matches on dob; unidentifiable!!! test against newer grades data

# update the temp$idnum_clean
describe(temp$idnum_clean)
temp$idnum_clean <- with(temp, ifelse(!is.na(id_verif), id_verif, idnum_clean))
temp$idnum_clean <- with(temp, ifelse(!is.na(init_verif), init_verif, idnum_clean))
temp$idnum_clean <- with(temp, ifelse(!is.na(judg_verif), judg_verif, idnum_clean))

# update your master data file
    temp$idnum_clean %in% f2$idnum_clean # none of these idnum cleans were in already in f2.
    temp_fm <- subset(temp, !is.na(idnum_clean), select = c(idnum, idnum_clean))
    sum(duplicated(temp_fm$idnum)) #2
    subset(temp_fm, duplicated(idnum) == T) # 38430
      # how many times is 38430 in your data? once!
      subset(f2, idnum == "38430", select = dob)
      # you need to pick who 38430 really is? 

```

### Update Oct2015 data for student withdrawals
```{r}
setwd("~/Box Sync/LD/10_LD_data_merging/07_Oct2015_LDsurvey")
ms <- read.csv('nontakersOct2015_TL.csv', head = T)
dim(ms)
names(ms)
ms$withdrewBy2015 <- NA
ms$withdrewBy2015 [ms$Withdrawn. == "Y"] <- 1
    ms$Withdrawn. <- NULL
summary(ms$withdrewBy2015)

summary(ms$Absent.) 
ms$absentOct2015 <- NA
ms$absentOct2015 [ms$Absent. == "Y" | ms$Absent. == "A" | ms$Absent. == "ALL DAY"] <- 1
summary(ms$absentOct2015); table(ms$absentOct2015)
ms$Absent. <- NULL

ms$SrAlreadyTookEcon == NA
ms$SrAlreadyTookEcon [ms$Senior...Already.took.Economics == "Y"] <- "Y"
ms$SrAlreadyTookEcon [ms$Senior...Already.took.Economics == "N"] <- "N"
ms$SrAlreadyTookEcon
ms$Senior...Already.took.Economics <- NULL

ms$Grade1415 <- NULL
ms$Grade1516 <- NULL

# dataset ms is ready for merging with master data
names(ms)
```
###!!!So this chunk ends and I have a list of withdrawn students to drop, but not yet a sense of how to drop them!!!

# VI. Fall 2015 Performance data
```{r Load Fall 2015 performance data}
setwd("~/Box Sync/LD/10_LD_data_merging/08_Fall_2015 Performance Data")

gort_f1516 <- read.csv("08_GORT_Fall1516.csv", 
               na.strings = "", # ensures empty cells of factor and character variables are read as NAs not "". 
               stringsAsFactors=FALSE,  # helps in some field data
               head = T) 

grades_f1516 <- read.csv("08_gradesFall1516.csv", head = T) 

# clean up the new data
names(gort_f1516)
    # rename StudentID_Transform as idnum clean, but this is temporary!!! It's not been vetted !!!
    gort_f1516$idnum_clean <- gort_f1516$StudentID_Transform; gort_f1516$StudentID_Transform <- NULL
    gort_f1516 <- moveme(gort_f1516, c("idnum_clean"), "first")
    # dropping lots of variables for short-term test
    gort_f1516 <- subset(gort_f1516, select = -c(Pattern, DSM_1:WISC_ProcessingSpeed))
    # suffix label all yor variables
    names(gort_f1516)[2:length(names(gort_f1516))] <- paste0(names(gort_f1516)[2:length(names(gort_f1516))], "_f1516")

# merge into m
names(m)
names(gort_f1516)
m2 <- merge(m, gort_f1516, by = c("idnum_clean"), all = T)

dim(m)
dim(gort_f1516)
dim(m2)
m <- m2

# merge in the grades data now
# clean up the new data
names(grades_f1516)
    # rename StudentID_Transform as idnum clean, but this is temporary!!! It's not been vetted !!!
    grades_f1516$idnum_clean <- grades_f1516$Student_ID_Transform; grades_f1516$Student_ID_Transform <- NULL
    grades_f1516 <- moveme(grades_f1516, "idnum_clean", "first")
    

# transform long grades data
sum(is.na(grades_f1516$Course)) # 0 

# Prepare grades data for merging
# paste course identyifing data into one variable course_yr_term
    grades_f1516$Course2 <- gsub("[ ]", "_", grades_f1516$Course)
    grades_f1516$course_yr_trm <- paste(grades_f1516$Course2, "_1516t", grades_f1516$Semester_Term, sep="")
    grades_f1516$course_yr_trm <- gsub("[:]", "", grades_f1516$course_yr_trm)

# rename performance variable
grades_f1516$score <- grades_f1516$Grade; grades_f1516$Grade <- NULL

# Convert data from long to wide format using cast()
  # create subset that drops nameless courses from data
  cp <- subset(grades_f1516, !is.na(Course))
  dim(cp) # 3083
  sum(is.na(cp$idnum_clean)) # 0 NAs

  # view key variables
  describe(cp$idnum_clean); length(unique(cp$idnum_clean)) # 472 unique IDs
  summary(cp$course_yr_trm)
  describe(cp$score)
  
  # create a subset of the variables you will transform to a wide dataset
  cp2 <- subset(cp, select = c(idnum_clean, course_yr_trm, score))
  dim(cp2)
  head(cp2)

# load packages
library(reshape2)
w <- dcast(cp2, idnum_clean ~ course_yr_trm,
           value.var = "score",
           fun.aggregate = mean, 
           na.rm = TRUE)
      # what is the effect of na.rm=T here???
      # what is the effect of mean in fun.aggregate???
      # how can I confirm if any data were lost???

dim(w) 
  # 472 unique idnumber matches the number of unique IDs in  describe(cp$idnum_clean)
  # 167 + 1 (ID col) columns matches describe(cp$course_yr_trm) values for unique course_yr_trm values
# test <- write.csv(w, 'test.csv') 

# merge in these grades with your data
m2 <- merge(m, w, by = "idnum_clean", all.x = T)
dim(m); dim(w); dim(m2)
d <- m2
```
# VIII. April 2016 Survey Data

```{r load april 2016 data}
setwd("~/Box Sync/LD/10_LD_data_merging")

# bring your custom functions in
source("00_LDMS_data_prep_package.R")

# load the spring survey 2016 (ss16)
t <- read.csv("09_Apr2016survey/09.1_LD_Mindset_04.29.16.csv", 
              na.strings = "", 
              stringsAsFactors=FALSE, 
              head = T) 

dim(t)
str(t)

t_desc <- t[1,] # save description row
t <- t[-1,] # get rid of description row

# rename your rownames b/c you dropped a row
  rownames(t) <- seq(length=nrow(t))
  
# Fix up any other weird things in the data here
t <- smart_cast_numeric(t)
class(t)
# d_desc["item_name"]

names(t)

t$Qual_ID <- t$V1; V1 <- NULL

t <- moveme(t, "Qual_ID", "first")
```

```{r drop your timing variables}
# display sorted names so you can ID the timing variables
sort(names(t))

t <- subset(t, select = -c(
  mp3tm_1, mp3tm_2, 
  mp3tm_3, mp3tm_4, 
  Q112_1, Q112_2, 
  Q112_3, Q112_4, 
  Q113_1, Q113_2, 
  Q113_3, Q113_4, 
  Q114_1, Q114_2, 
  Q114_3, Q114_4, 
  Q115_1, Q115_2, 
  Q115_3, Q115_4, 
  Q116_1, Q116_2, 
  Q116_3, Q116_4, 
  Q118_1, Q118_2, 
  Q118_3, Q118_4, 
  Q120_1, Q120_2, 
  Q120_3, Q120_4, 
  Q173_1, Q173_2, 
  Q173_3, Q173_4, 
  Q174_1, Q174_2, 
  Q174_3, Q174_4, 
  Q175_1, Q175_2, 
  Q175_3, Q175_4, 
  Q216, Q217))

t <- subset(t, select = -c(
  Q218, Q219_1, 
  Q219_2, Q219_3, 
  Q219_4, Q221_1, 
  Q221_2, Q221_3, 
  Q221_4, Q222_1, 
  Q222_2, Q222_3, 
  Q222_4, Q223_1, 
  Q223_2, Q223_3, 
  Q223_4, Q224_1, 
  Q224_2, Q224_3, 
  Q224_4, Q225_1, 
  Q225_2, Q225_3, 
  Q225_4, Q226_1, 
  Q226_2, Q226_3, 
  Q226_4, Q227_1, 
  Q227_2, Q227_3, 
  Q227_4, Q265_1, 
  Q265_2, Q265_3, 
  Q265_4, Q266_1, 
  Q266_2, Q266_3, 
  Q266_4, Q267_1, 
  Q267_2, Q267_3, 
  Q267_4, Q268_1, 
  Q268_2, Q268_3, 
  Q268_4, Q269_1))

t <- subset(t, select = -c(
  Q269_2, Q269_3, 
  Q269_4, Q270_1, 
  Q270_2, Q270_3, 
  Q270_4, Q271_1, 
  Q271_1.1, Q271_2, 
  Q271_2.1, Q271_3, 
  Q271_3.1, Q271_4, 
  Q271_4.1, Q272_1, 
  Q272_2, Q272_3, 
  Q272_4, Q273_1, 
  Q273_2, Q273_3, 
  Q273_4, Q274_1, 
  Q274_2, Q274_3, 
  Q274_4, Q275_1, 
  Q275_2, Q275_3, 
  Q275_4, Q276_1, 
  Q276_2, Q276_3, 
  Q276_4, Q278_1, 
  Q278_2, Q278_3, 
  Q278_4, Q280_1, 
  Q280_2, Q280_3, 
  Q280_4, Q281_1, 
  Q281_2, Q281_3, 
  Q281_4, Q282_1, 
  Q282_2, Q282_3, 
  Q282_4, Q283_1, 
  Q283_2, Q283_3, 
  Q283_4, Q284_1, 
  Q284_2, Q284_3, 
  Q284_4, Q286_1))

t <- subset(t, select = -c(
  Q286_2, Q286_3, 
  Q286_4, Q288_1, 
  Q288_2, Q288_3, 
  Q288_4, Q291_1, 
  Q291_2, Q291_3, 
  Q291_4, Q292_1, 
  Q292_2, Q292_3, 
  Q292_4, Q293_1, 
  Q293_2, Q293_3, 
  Q293_4, Q294_1, 
  Q294_2, Q294_3, 
  Q294_4 
  ))

# drop other variables
names(t) %>% noquote()

t <- subset(t, select = -c(V1:V7, IDS, Intro1))

t <- subset(t, select = -c(LocationLatitude, LocationLongitude, LocationAccuracy, X, V8, V9, V10))
```

### clean up april 2016 survey idnumbers
```{r clean up april 2016 survey idnumbers}
# sort idnums to display idnums corrupted with text
sort(t$idnum)

# save an original copy of idnumbers
t$idnum.orig <- t$idnum

# repair ids with text in them
class(t$idnum)
t$idnum [t$idnum == "__________19240"] <-  "19240"      
t$idnum [t$idnum == "__________42600"] <- "42600"          
t$idnum [t$idnum == "__________33260"] <- "33260"           
t$idnum [t$idnum == "__________1950"] <- "19500"                        
t$idnum [t$idnum == "__________5401"] <- "54010" 
t$idnum [t$idnum == "__________4443"] <- "44430"          
t$idnum [t$idnum == "__________0"] <- NA   
t$idnum [t$idnum == "__________16"] <- NA

# add a 0 to 4-digit idnumbers
t$idnum <- ifelse(nchar(t$idnum) == 4, paste(t$idnum, 0, sep = ""), t$idnum)

# test if variable contains any letters of alphabet
t$idnum %>% grepl('^[A-Za-z]+$', .) %>% table
      # means check from the start (^) to end (+$) 
      # of string if strong contains A-Z (caps) or a-z (lowercase) 
      # and return TRUE/FALSE. 

# display if variable contains only numbers --> discovery of one more case with invalid ID number. 
t %>% subset( grepl('^[0-9]+$', idnum) == FALSE, select = idnum)
    # reveals row 52 wrote "4#'s0" for his idnumber
    t[51,] # has intials MEB, dob 03/21/02, grade 8. Find him in the school records!!!

# reclassify cleaned ids as numeric
t$idnum <- as.numeric(t$idnum)
```

## Confirm ID numbers entered by students against a school-provided key
### How many cases are we starting with? (418)
```{r}
names(t)
t <- moveme(t, "idnum.orig", "after", "idnum")

# reclass idnum and initials as character vars
class(t$idnum)
t$idnum <- as.character(t$idnum)
class(t$initials)

dim(t)
```

### load a school-provided key 
```{r load Spring 2016 grades data}
# load the Spring 2016 grades data
s <- read.csv("10_Spring_2016_grades/ld_spr_2016_grades.csv", 
              na.strings = "", 
              stringsAsFactors=FALSE, 
              head = T) 

s_desc <- s[1,] # save description row
s <- s[-1,] # get rid of description row

# rename your rownames b/c you dropped a row
  rownames(s) <- seq(length=nrow(s))
  
# Fix up any other weird things in the data here
s <- smart_cast_numeric(s)
class(s)
# d_desc["item_name"]

names(s)
```

```{r resolve survey IDs that don't match School provided grades}
# identify a school-provided key of legit IDnumbers from the Spring 2016 grades
s$Student_ID_Transform

# Check IDs for duplicates
subset(t, duplicated(idnum) == TRUE, select = idnum)

# Correct your duplicate cases: 
subset(t, idnum == '18640') # 2 different people
  # per 03_Oct2014_Student_Initials_ID, 18640 has initials JC, which matches 2nd row
  # 33160, 33180, 18780 have AG initials When you find your school provided dob's you can ID her. Temporarily drop!!!
  t <- subset(t, Qual_ID != 'R_XmSbBh99wkiDalH') 
subset(t, idnum == '32960') # 1 person, 2nd entry is blank
  t <- subset(t, Qual_ID != "R_suRUlDN55d5imgV") # drop blank row
subset(t, idnum == '40040') # same initials & dob, second row is blank
  t <- subset(t, Qual_ID != "R_2B4dlASyv5bSfAT") # drop blank row
subset(t, idnum == '33120') # second row is blank
  t <- subset(t, Qual_ID != "R_6wW7VAfvwrXjB29") # drop blank row
  
t <- subset(t, !is.na(idnum)) # Drop your NA IDs temporarily!!!

dim(t) # so now I have 285 clean Ps from the Spring 2016 survey
``` 

### merge t with d
```{r}
names(t)
t <- moveme(t, 'idnum', 'before', 'Qual_ID')
names(t)[2:length(names(t))] <- paste0(names(t)[2:length(names(t))], "_s16") # 2: means to start with the second name

names(t)
names(d)

class(d$idnum_clean)
class(t$idnum)

t$idnum %>% duplicated %>% table
d$idnum_clean %>% duplicated %>% table
```

#### temp resolve 16 duplicates in d !!!
```{r}
subset(d, duplicated(idnum_clean) == TRUE, select = idnum_clean)
d <- subset(d, !is.na(idnum)) # Drop your NA IDs temporarily!!!

# may be caused by multiple tests in some cases
subset(d, idnum_clean == '19590') # all info is redundant --> drop second row
subset(d, idnum_clean == '19610') # all info is redundant --> drop second row
subset(d, idnum_clean == '19660') # check closer
subset(d, idnum_clean == '25910') #
subset(d, idnum_clean == '32890') #
subset(d, idnum_clean == '33410') #
subset(d, idnum_clean == '34460') #
subset(d, idnum_clean == '35820') #
subset(d, idnum_clean == '35830') #
subset(d, idnum_clean == '35970') #
subset(d, idnum_clean == '36230') #
subset(d, idnum_clean == '39970') #
subset(d, idnum_clean == '41500') #

# Temporarily dropping all idnum_cleans that are dups!!!
dim(d)
d <- subset(d, duplicated(idnum_clean) == FALSE)
```

```{r merger of t with d}
# Now there are no dups
t$idnum %>% duplicated %>% table
d$idnum_clean %>% duplicated %>% table

names(t)

dim(d); dim(t)

d <- merge(d, t, 
      by.x = "idnum_clean", 
      by.y = "idnum", 
      all = T)

dim(d)
```

```{r}
t %>% names %>% noquote

t_desc['sm_better1']
t_desc['sm_help']   
t_desc['sm_better2'] 
t_desc['sm_stkhelp'] 
t_desc['sm_bttrway'] 
t_desc['sm_dobett']
t_desc['SM1_r']      
t_desc['SM2'] 

# These are timing variables
t_desc['sm_dotm_1']  
t_desc['sm_dotm_2']  
t_desc['sm_dotm_3']  
t_desc['sm_dotm_4'] 
```

# IX. Spring 2016 Performance Data 
```{r}
# already loaded in a previous chunk
# load the Spring 2016 grades data
# s <- read.csv("10_Spring_2016_grades/ld_spr_2016_grades.csv", 
#              na.strings = "", 
#              stringsAsFactors=FALSE, 
#              head = T) 

# Make your long grades data wide
names(s)
names(s) [names(s) == 'Student_ID_Transform'] <- 'idnum'

sum(is.na(s$Course)) # no NAs for Course titles 

# Prepare grades data for merging
# paste course identyifing data into one variable course_yr_term
    s$Course <- gsub("[ ]", "_", s$Course) # replace " " with "_"
    s$Course <- gsub("[:]", "_", s$Course) # replace " " 
    s$Course <- gsub("[-]", "_", s$Course) # replace " " 
    s$course_yr_trm <- with(s, paste(Course, "_1516t", Semester_Term, sep=""))
# rename performance variable
    s$score <- s$Grade; s$Grade <- NULL

# Convert data from long to wide format using cast()
  # create subset that drops nameless courses from data
  cp <- s
  dim(cp)
  sum(is.na(cp$idnum)) # 0 NAs

  # view key variables
  length(unique(cp$idnum)) # 573 unique IDs
  
  # create a subset of the variables you will transform to a wide dataset
  cp2 <- subset(cp, select = c(idnum, course_yr_trm, score))
  dim(cp2)
  head(cp2)

# load packages
library(reshape2)
w <- reshape2::dcast(cp2, idnum ~ course_yr_trm,
                   value.var = "score",
                   fun.aggregate = mean, 
                   na.rm = TRUE)
                  # what is the effect of na.rm=T here???
                  # what is the effect of mean in fun.aggregate???
                  # how can I confirm if any data were lost???

dim(w) 
head(w)
```

### You should now be able to merge the wide grades data (w) with the main dataset (d)
```{r}
names(w)

names(d)

# confirm key vars are of same class
class(d$idnum_clean)
class(w$idnum)
w$idnum <- as.character(w$idnum)

# Now there are no dups
w$idnum %>% duplicated %>% table
d$idnum_clean %>% duplicated %>% table

dim(d); dim(w)

d <- merge(d, w, 
      by.x = "idnum_clean", 
      by.y = "idnum", 
      all = T)

dim(d) # looks like only 183 of original Ps made it through
```

```{r}
############ Resolve Algebra ############ 
subset(d, select = c(Algebra_I_1516t2.x,  Algebra_I_1516t2.y) )
d$Algebra_I_1516t2 <- with(d, paste(Algebra_I_1516t2.x, Algebra_I_1516t2.y, sep=''))
d$Algebra_I_1516t2 <- d$Algebra_I_1516t2 %>% str_replace_all("NaN", "") %>% str_replace_all("NA", "") %>% as.numeric
table(d$Algebra_I_1516t2) # looks clean

# drop .x & .y versions
d <- subset(d, select = -c(Algebra_I_1516t2.x, Algebra_I_1516t2.y))
names(d)


########### Resolve Spanish ############ 
subset(d, select = c(Spanish_I_1516t2.x,  Spanish_I_1516t2.y) )
d$Spanish_I_1516t2 <- with(d, paste(Spanish_I_1516t2.x, Spanish_I_1516t2.y, sep=''))
d$Spanish_I_1516t2 <- d$Spanish_I_1516t2.y %>% str_replace_all("NaN", "") %>% str_replace_all("NA", "") %>% as.numeric
table(d$Spanish_I_1516t2) # looks clean

# drop .x & .y versions
d <- subset(d, select = -c(Spanish_I_1516t2.x, Spanish_I_1516t2.y))
```

```{r add grade level from official roster of 2015-16 to main data}
# create a subset of the variables you will transform to a wide dataset
  cp3 <- subset(cp, select = c(idnum, Grade_In_School))
  dim(cp3)
  head(cp3)

# creates a dataframe with cols 6-12 
g <- reshape2::dcast(cp3, idnum ~ Grade_In_School,
                   value.var = "Grade_In_School",
                   fun.aggregate = mean, 
                   na.rm = TRUE)
# rename vars 
names(g) [names(g) == '6']  <- "six"
names(g) [names(g) == '7']  <- "seven"
names(g) [names(g) == '8']  <- "eight"
names(g) [names(g) == '9']  <- "nine"
names(g) [names(g) == '10'] <- "ten"
names(g) [names(g) == '11'] <- "eleven"
names(g) [names(g) == '12'] <- "twelve"

# condense
g$official_gr_1516t2 <- apply(X = g[,c('six', 'seven', 'eight', 'nine', 'ten', 'eleven', 'twelve')],
                MARGIN =1, FUN = mean, na.rm = T)
head(g)

# reduce data to vars you wish to merge
g <- subset(g, select=c(idnum, official_gr_1516t2))

dim(d); dim(g)

d <- merge(d, g, 
      by.x = "idnum_clean", 
      by.y = "idnum", 
      all = T)

dim(d)
```

## Correct var names that begin with a number
```{r}
# write the correct names pseudofunction o/s the denumber_names function so it can be applied to cell contents as well as colnames. 
correct_names <- . %>% 
    str_replace("6th", "Sixth") %>% 
    str_replace("7th", "Seventh") %>% 
    str_replace("8th", "Eighth")

denumber_names <- function(x) {
  cn <- colnames(x) # save list of column names as cn
  cn <- correct_names(cn) # apply renaming function to list of column names
  colnames(x) <- cn # overwrite summary stats table with clean names
  x
}

d <- denumber_names(d)
```

### Create a long dataframe of average grade levels per course (gr_lev_long)
```{r convert array to dataframe}
cp4 <- subset(cp, select = c(course_yr_trm, Grade_In_School))
  dim(cp4)
  head(cp4)
  
gr_lev <- base::tapply(cp4$Grade_In_School, cp4$course_yr_trm, mean) 
    # - creates an array 
    # - dimnames(gr_lev) indicates the factor level of course_yr_trm

gr_lev_long <- data.frame(gr_lev, rownames = dimnames(gr_lev))
    # Drop weird 2nd variable created during dataframe conversion
    gr_lev_long <- subset(gr_lev_long, select = gr_lev)

# create new var to indicate the course
gr_lev_long$Course <- NA
gr_lev_long$Course <- rownames(gr_lev_long)
# reset row names
  rownames(gr_lev_long) <- seq(length=nrow(gr_lev_long))
  
gr_lev_long$Course <- gr_lev_long$Course %>% correct_names
```


These data could still be more thoroughly checked for 
!!!
???
and missing Ps


## 2013-2014 Spring GPAs
```{r composite Math grade 1314t1}
# create a math composite variable
d$math_cv_1314t1 <- apply(X = d[,c('Algebra_I_-_Modified_1314t1', 
                      'Algebra_I_1314t1', 
                      'Algebra_II_-_Honors_1314t1', 
                      'Algebra_II_1314t1', 
                      'Economics_1314t1', 
                      'Geometry_-_Honors_1314t1', 
                      'Geometry_-_Modified_1314t1', 
                      'Geometry_1314t1', 
                      'Independent_Study_in_Math_-_College_Algebra_1314t1', 
                      'Math_Models_1314t1', 
                      'Math_Models_Modified_1314t1', 
                      'Pre-Calculus_-_Honors_1314t1')], 
                      MARGIN = 1, FUN = mean, na.rm = T)
hist(d$math_cv_1314t1)
      
describe(d$math_cv_1314t1)

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$math_cv_1314t1)
hist(s16$math_cv_1314t1)
stem(s16$math_cv_1314t1)
```

```{r composite Science grade 1314t1}
d$science_cv_1314t1 <- apply(X = d[,c('Anatomy_1314t1', 
                            'Biology_-_Honors_1314t1', 
                            'Biology_-_Modified_1314t1', 
                            'Biology_1314t1', 
                            'Chemistry_-_Honors_1314t1', 
                            'Chemistry_-_Modified_1314t1', 
                            'Chemistry_1314t1')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$science_cv_1314t1)
      
describe(d$science_cv_1314t1) # big skew, but N = 543

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$science_cv_1314t1)
hist(s16$science_cv_1314t1)
stem(s16$science_cv_1314t1)
```

```{r composite English grade 1314t1}
d$english_cv_1314t1 <- apply(X = d[,c('English_I_-_Honors_1314t1', 
                            'English_I_1314t1', 
                            'English_II_-_Honors_1314t1', 
                            'English_II_-_Modified_1314t1', 
                            'English_II_1314t1', 
                            'English_III_-_American_Lit_-_Honors_1314t1', 
                            'English_III_-_American_Lit_1314t1')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$english_cv_1314t1)
describe(d$english_cv_1314t1) 

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$english_cv_1314t1)
hist(s16$english_cv_1314t1)
stem(s16$english_cv_1314t1)
```

```{r composite Reading grade 1314t1}
d$reading_cv_1314t1 <- apply(X = d[,c('Reading_-_SARA__III_1314t1', 
                            'Reading_-_SARA_I_1314t1', 
                            'Reading_-_SARA_II_1314t1', 
                            'Reading_-_SARA_III_1314t1', 
                            'Reading_Comprehension_I_1314t1', 
                            'Reading_Comprehension_II_1314t1', 
                            'Reading_Comprehension_III_1314t1', 
                            'Reading_SARA_-_III_1314t1')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$reading_cv_1314t1)
      
describe(d$reading_cv_1314t1) 

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$reading_cv_1314t1)
hist(s16$reading_cv_1314t1)
stem(s16$reading_cv_1314t1)
```

```{r core subject GPA 1314t1}
d$core_gpa_1314t1 <- apply(X = d[,c('math_cv_1314t1',
                            'science_cv_1314t1',
                            'english_cv_1314t1',
                            'reading_cv_1314t1')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$core_gpa_1314t1)
      
describe(d$core_gpa_1314t1) 

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$core_gpa_1314t1)
hist(s16$core_gpa_1314t1)
stem(s16$core_gpa_1314t1)
```

## 2015-2016 Spring GPAs
```{r composite Math grade 1516t2}
# create a math composite variable
d$math_cv_1516t2 <- apply(X = d[,c('Math_Models_1516t2', 'Pre_Calculus___Honors_1516t2', 'Calculus___Honors_1516t2', 'Geometry___Honors_1516t2', 'Geometry_1516t2', 'Algebra_I_1516t2', 'Algebra_II___Honors_1516t2', 'Algebra_II___Modified_1516t2', 'Algebra_II_1516t2', 'Eighth_Math__Algebra_I_1516t2', 'Eighth_Grade_Mathematics_1516t2', 'Seventh_grade_Mathematics_1516t2', 'Sixth_Mathematics_1516t2')], 
                  MARGIN = 1, FUN = mean, na.rm = T)
hist(d$math_cv_1516t2)
      
describe(d$math_cv_1516t2) # big skew, but N = 543

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$math_cv_1516t2)
hist(s16$math_cv_1516t2)
stem(s16$math_cv_1516t2)
```

```{r composite Science grade 1516t2}
d$science_cv_1516t2 <- apply(X = d[,c('Sixth_Science_1516t2',
                            'Seventh_Science_1516t2',
                            'Eighth_Science_1516t2',
                            'Anatomy_1516t2',
                            'Biology___Honors_1516t2',
                            'Biology_1516t2',
                            'Chemistry___Honors_1516t2',
                            'Chemistry_1516t2',
                            'Earth_and_Space_Science_1516t2',
                            'Physics_I___Honors_1516t2',
                            'Physics_I_1516t2',
                            'Physics_II___Honors_1516t2')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$science_cv_1516t2)
      
describe(d$science_cv_1516t2) # big skew, but N = 543

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$science_cv_1516t2)
hist(s16$science_cv_1516t2)
stem(s16$science_cv_1516t2)
```

```{r composite English grade 1516t2}
d$english_cv_1516t2 <- apply(X = d[,c('Sixth_English___Assoc._Method_1516t2',                               
                            'Sixth_English_1516t2',
                            'Seventh_English_1516t2',
                            'Eighth_English_1516t2',
                            'English_I___Honors_1516t2',
                            'English_I_1516t2',
                            'English_II___Honors_1516t2',
                            'English_II_1516t2',
                            'English_III___American_Lit___Honors_1516t2',
                            'English_III___American_Lit_1516t2',
                            'English_IV___British_Lit___Honors_1516t2',
                            'English_IV___British_Lit_1516t2')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$english_cv_1516t2)
describe(d$english_cv_1516t2) 

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$english_cv_1516t2)
hist(s16$english_cv_1516t2)
stem(s16$english_cv_1516t2)
```

```{r composite Reading grade 1516t2}
d$reading_cv_1516t2 <- apply(X = d[,c('Sixth_Reading_1516t2',
                            'Seventh_English___Association_Method_1516t2',
                            'Seventh_Reading_1516t2',
                            'Eighth_Reading_1516t2',
                            'Reading___10_1516t2',
                            'Reading___11_1516t2',
                            'Reading___9_1516t2')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$reading_cv_1516t2)
      
describe(d$reading_cv_1516t2) 

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$reading_cv_1516t2)
hist(s16$reading_cv_1516t2)
stem(s16$reading_cv_1516t2)
```

```{r core subject GPA 1516t2}
d$core_gpa_1516t2 <- apply(X = d[,c('math_cv_1516t2',
                            'science_cv_1516t2',
                            'english_cv_1516t2',
                            'reading_cv_1516t2')], 
                            MARGIN = 1, FUN = mean, na.rm = T)

hist(d$core_gpa_1516t2)
      
describe(d$core_gpa_1516t2) 

# subset who completed a Strategic Mindset Scale (N=265)
s16 <- subset(d, !is.na(strat_ms_cv))
describe(s16$core_gpa_1516t2)
hist(s16$core_gpa_1516t2)
stem(s16$core_gpa_1516t2)
```

# Composite variable calculation for all studies:
```{r CV: GORT}
# GORT 
d$gort_1516 <- apply(X = d[,c("RA_f1516", "RC_f1516","RR_f1516")],
                MARGIN =1, FUN = mean, na.rm = T)
      d <- moveme(d, c("gort_1516"), "after", "GORT_TestDate_f1516")
      label(d$gort_1516) = "mean GORT score from 2015-16 data" 
      describe(d$gort_1516)
      hist(d$gort_1516)
```

```{r CV: mindset}
########################################################################################          
######################### Growth Mindset Fall 2014-15 ##################################
########################################################################################

    d$ITI_amount_f14_R <- 7 - d$ITI_amount_f14 # "You have a certain amount of intelligence and you really can`t do much to change  it."
    d$ITI_change_f14_R <- 7 - d$ITI_change_f14 # "Your intelligence is something about you that you can`t change very  much."
    d$ITI_basic_f14_R <- 7 - d$ITI_basic_f14 # "You can learn new things, but you can`t really change your basic intelligence."
    
    d$ ITI_comp_f14 <- apply(X = d[,c("ITI_amount_f14_R","ITI_change_f14_R", "ITI_basic_f14_R")], 
                              MARGIN =1, FUN = mean, na.rm = T)
      
    hist(d$ ITI_comp_f14)
    pd <- density(d$ ITI_comp_f14, na.rm=T); plot(pd) 
    describe(d$ ITI_comp_f14)
    
          temp <- subset(d, select = c("ITI_amount_f14_R","ITI_change_f14_R", "ITI_basic_f14_R"))
          # get the column numbers for the variables
          variables = names(temp); as.data.frame(variables) 
          # factor analyze your variables
          factanal(na.omit(temp[,1:3]),1) # the load
          alpha(na.omit(temp[,c(1:3)])) # a = .

########################################################################################          
######################### Growth Mindset Spring 2015 ###################################
########################################################################################
          
    d$ITI_amount_s15_R <- 7 - d$ITI_amount_s15 # "You have a certain  amount of intelligence and you really can`t do much to change  it."
    d$ITI_change_s15_R <- 7 - d$ITI_change_s15 # "Your intelligence is  something about you that you can`t change very  much."
    d$ITI_basic_s15_R <- 7 - d$ITI_basic_s15 # "You can learn new  things, but you can`t really change your basic intelligence."
    
    d$ ITI_comp_s15 <- apply(X = d[,c("ITI_amount_s15_R","ITI_change_s15_R", "ITI_basic_s15_R")], 
                              MARGIN =1, FUN = mean, na.rm = T)
      
    hist(d$ ITI_comp_s15)
    pd <- density(d$ ITI_comp_s15, na.rm=T); plot(pd) 
    describe(d$ ITI_comp_s15)
    
          temp <- subset(d, select = c("ITI_amount_s15_R","ITI_change_s15_R", "ITI_basic_s15_R"))
          # get the column numbers for the variables
          variables = names(temp); as.data.frame(variables) 
          # factor analyze your variables
          factanal(na.omit(temp[,1:3]),1) # the load
          alpha(na.omit(temp[,c(1:3)])) # a = .

########################################################################################          
######################### Growth Mindset Fall 2015 #####################################
########################################################################################

    d$ITI_amount_f15_R <- 7 - d$ITI_amount_f15 # "You have a certain  amount of intelligence and you really can`t do much to change  it."
    d$ITI_change_f15_R <- 7 - d$ITI_change_f15 # "Your intelligence is  something about you that you can`t change very  much."
    d$ITI_basic_f15_R <- 7 - d$ITI_basic_f15 # "You can learn new  things, but you can`t really change your basic intelligence."
    
    d$ ITI_comp_f15 <- apply(X = d[,c("ITI_amount_f15_R","ITI_change_f15_R", "ITI_basic_f15_R")], 
                              MARGIN =1, FUN = mean, na.rm = T)
      
    hist(d$ ITI_comp_f15)
    pd <- density(d$ ITI_comp_f15, na.rm=T); plot(pd) 
    describe(d$ ITI_comp_f15)
    
          temp <- subset(d, select = c("ITI_amount_f15_R","ITI_change_f15_R", "ITI_basic_f15_R"))
          # get the column numbers for the variables
          variables = names(temp); as.data.frame(variables) 
          # factor analyze your variables
          factanal(na.omit(temp[,1:3]),1) # they load
          alpha(na.omit(temp[,c(1:3)])) # a = .81

########################################################################################          
######################### Growth Mindset Spring 2016 ###################################
########################################################################################
          
    d$ITI_amount_s16_R <- 7 - d$ITI_amount_s16 # "You have a certain  amount of intelligence and you really can`t do much to change  it."
    d$ITI_change_s16_R <- 7 - d$ITI_change_s16 # "Your intelligence is  something about you that you can`t change very  much."
    d$ITI_basic_s16_R <- 7 - d$ITI_basic_s16 # "You can learn new  things, but you can`t really change your basic intelligence."
    
    d$ITI_comp_s16 <- apply(X = d[,c("ITI_amount_s16_R","ITI_change_s16_R", "ITI_basic_s16_R")], 
                              MARGIN =1, FUN = mean, na.rm = T)
    
    d <- moveme(d, 'ITI_comp_s16', 'after', 'ITI_amount_s16')
    d <- moveme(d, c("ITI_amount_s16_R", "ITI_change_s16_R", "ITI_basic_s16_R"), 'after', 'ITI_comp_s16')
      
    hist(d$ITI_comp_s16)
    pd <- density(d$ITI_comp_s16, na.rm=T); plot(pd) 
    describe(d$ITI_comp_s16)
    
          temp <- subset(d, select = c("ITI_amount_s16_R","ITI_change_s16_R", "ITI_basic_s16_R"))
          # get the column numbers for the variables
          variables = names(temp); as.data.frame(variables) 
          # factor analyze your variables
          factanal(na.omit(temp[,1:3]),1) # they load
          alpha(na.omit(temp[,c(1:3)])) # a = .83
```         
          
```{r experimental condition}
# contrast coded mindset condition
d$cond_ms_cc <- NA
d$cond_ms_cc [d$cond_ms_s15 == "Control"] <- -1
d$cond_ms_cc [d$cond_ms_s15 == "Growth"] <- 1
d <- moveme(d, c("cond_ms_cc"), "after", "cond_ms_s15") 

# mindset condition coded 12
d$cond_ms_n <- NA
d$cond_ms_n [d$cond_ms_s15 == "Control"] <- 1
d$cond_ms_n [d$cond_ms_s15 == "Growth"] <- 2
d <- moveme(d, c("cond_ms_n"), "after", "cond_ms_s15") 

# mindset difference score
d$ITI_diff_f14_to_f15 <- d$ITI_comp_f15 - d$ITI_comp_f14
hist(d$ITI_diff_f14_to_f15)
d <- moveme(d, c("ITI_diff_f14_to_f15"), "after", "ITI_comp_f15") 
```

```{r IQ & GORT}
# WISC IQ
d$WISC_1415 <- apply(X = d[,c("WorkingMemory_1415", "VerbalComp_1415", "PerceptualReasoning_1415", "ProcessingSpeed_1415")], MARGIN =1, FUN = mean, na.rm = T)

# GORT
d$gort_1415 <- apply(X = d[,c("GORT_RA_1415",  "GORT_RC_1415",  "GORT_RR_1415")], MARGIN =1, FUN = mean, na.rm = T)
```

## Strategic Behaviors
This strategic behaviors scale may be too low alpha to use meaningfully. 
```{r strategic behaviors}
# Fall 2014 strategic behaviors: 
      d$B_SP2_R_f14 <- 6 - d$B_SP2_f14
      d$B_SM7_R_f14 <- 6 - d$B_SM7r_f14
    # subset variables to test
    fact <- subset(d, select = c("B_SP1_f14",   # consider alts b4 jumping in
                                 "B_SP2_R_f14", # jump right in
                                 "B_SP3_f14",   # try teacher's recomended approach
                                 "B_SM4_f14",   # stuck --> step back
                                 "B_SM5_f14",   # struggle --> mimic students
                                 "B_SM6_f14",   # struggle --> reach out to students for alt paths
                                 "B_SM7_R_f14"))# hard to change approach
    
    corr_table <- round(cor(fact, use="pairwise.complete.obs"), digits=3); corr_table
    
    corrplot(corr_table, method = "circle") #plot matrix
        # The strongest correlation is b/w B_SM6_f14 & B_SM5_f14, r = .40. These are peer helpseeking items. 
    
    scree_plot <- scree(corr_table,factors=TRUE, pc=TRUE, main="Scree plot", hline=NULL, add=FALSE); scree_plot 
    
    # Begin Factor Analysis
    dim(fact); str(fact)
    factanal(na.omit(fact[,1:ncol(fact)]), 1, rotation="varimax")
    alpha(na.omit(fact[,c(1:ncol(fact))])) # .48 is very low
    
    # could call this item peer_help but alpha is still low
    alpha(na.omit(fact[,c(5:6)])) # .57
```

These strategic behavior variables load so poorly they don't seem worth including. 
```{r low alpha strategic behavior composites}
# Strategic Behaviors

######################################################### 
###################### Fall 2014 ######################## 
#########################################################       
      d$strat_beh_f14 <- apply(X = d[,c("B_SP1_f14", "B_SP2_R_f14", "B_SP3_f14", "B_SM4_f14", "B_SM5_f14", "B_SM6_f14", "B_SM7_R_f14")], MARGIN =1, FUN = mean, na.rm = T)

      d <- moveme(d, c("B_SP2_R_f14", "B_SM7_R_f14", "strat_beh_f14"), "after", "B_SM7r_f14")
            # SPSP_2016 poster just interacted B_SP_3
      
######################################################### 
###################### Spring 2015 ###################### 
#########################################################      
      d$B_SP2_R_s15 <- 6 - d$B_SP2_s15
      d$B_SM7_R_s15 <- 6 - d$B_SM7r_s15
      
      d$strat_beh_s15 <- apply(X = d[,c("B_SP1_s15", "B_SP2_R_s15", "B_SP3_s15", "B_SM4_s15", "B_SM5_s15", "B_SM6_s15", "B_SM7_R_s15")], MARGIN =1, FUN = mean, na.rm = T)
      
      d <- moveme(d, c("B_SP2_R_s15", "B_SM7_R_s15", "strat_beh_s15"), "after", "B_SM7r_s15")

######################################################### 
###################### Fall 2015 ######################## 
#########################################################      
      d$B_SP2_R_f15 <- 6 - d$B_SP2_f15
      d$B_SM7_R_f15 <- 6 - d$B_SM7r_f15
      
      d$strat_beh_f15 <- apply(X = d[,c("B_SP1_f15", "B_SP2_R_f15", "B_SP3_f15", "B_SM4_f15", "B_SM5_f15", "B_SM6_f15", "B_SM7_R_f15")], MARGIN =1, FUN = mean, na.rm = T)
      
      d <- moveme(d, c("B_SP2_R_f15", "B_SM7_R_f15", "strat_beh_f15"), "after", "B_SM7r_f15")

######################################################### 
###################### Spring 2016 ###################### 
######################################################### 
# reverse score variables that need it      
d$B_SP2_R_s16 <- 6 - d$B_SP2_s16
d$B_SM7_R_s16 <- 6 - d$B_SM7r_s16

# create subset of variables to factor analyze
df_t <- subset(d, select = c("B_SP1_s16", "B_SP2_R_s16", "B_SP3_s16", "B_SM4_s16", "B_SM5_s16", "B_SM6_s16", "B_SM7_R_s16"))
 
    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot

# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 

# Alpha value for composite variable    
alpha(na.omit(df_t[,c("B_SP1_s16", "B_SP2_R_s16", "B_SP3_s16", "B_SM4_s16", "B_SM5_s16", "B_SM6_s16", "B_SM7_R_s16")])) 

# Create composite variable
d$strat_beh_s16 <- apply(X = d[,c("B_SP1_s16", "B_SP2_R_s16", "B_SP3_s16", "B_SM4_s16", "B_SM5_s16", "B_SM6_s16", "B_SM7_R_s16")], MARGIN=1, FUN=mean, na.rm=TRUE)

# Update variable order in main dataset
d <- moveme(d, c("B_SP2_R_s16", "B_SM7_R_s16", "strat_beh_s16"), "after", "B_SM7r_s16")
```
If none of the strategic behaviors factor load, then why are you combining them into a single variable? 

### Augmented strategic behavior scale (*strat_beh_small_s16*)
Intended to raise alpha to sufficient level
```{r}
# create subset of variables to factor analyze
df_t <- subset(d, select = c("B_SP1_s16", "B_SP3_s16", "B_SM4_s16", 'SM2_s16', 'SM_scav_s16'))
 
    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot

# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 

# Alpha value for composite variable    
a <- alpha(na.omit(df_t[,c("B_SP1_s16", "B_SP3_s16", "B_SM4_s16", 'SM2_s16', 'SM_scav_s16')])) # .61

# Create composite variable
d$strat_beh_small_s16 <- apply(X = d[,c("B_SP1_s16", "B_SP3_s16", "B_SM4_s16", 'SM2_s16', 'SM_scav_s16')], MARGIN=1, FUN=mean, na.rm=TRUE)

# Update variable order in main dataset
d <- moveme(d, c("strat_beh_small_s16"), "before", "B_SP1_s16")
```
Alpha is just under threshold in the augmented scale for strategic behavior, ($\alpha$ = `r a$total$raw_alpha %>% myround(2)`).

alphas for 2-item autonomy are .63, .59 and .61. You could probably rely on either this or the best single item autonomy and test if there is any difference.
```{r autonomy}
###################################################################
###################### Fall 2014-2015 #############################
###################################################################
      
          # Factor Analysis
          d$SS_A9_R_f14 <- 6 - d$SS_A9_f14
          d$SS_A11_R_f14 <- 6 - d$SS_A11_f14
          fact <- subset(d, select = c(SS_A9_R_f14, SS_A10_f14, SS_A11_R_f14))
          dim(fact); str(fact)
          corr_table <- round(cor(fact, use="pairwise.complete.obs"), digits=3); corr_table
          factanal(na.omit(fact[,1:ncol(fact)]), 1, rotation="varimax")
          alpha(na.omit(d[,c("SS_A9_R_f14", "SS_A11_R_f14")])) # .63
      d$autonind_f14 <- apply(X = d[,c("SS_A9_R_f14", "SS_A11_R_f14")], 
                        MARGIN =1, FUN = mean, na.rm = T)
      d <- moveme(d, c("SS_A9_R_f14", "SS_A11_R_f14", "autonind_f14"), "after", "SS_A12_f14")

###################################################################
###################### Spring 2014-2015 ###########################
###################################################################
      
      d$SS_A9_R_s15 <- 6 - d$SS_A9_s15
      d$SS_A11_R_s15 <- 6 - d$SS_A11_s15
      fact <- subset(d, select = c(SS_A9_R_s15, SS_A10_s15, SS_A11_R_s15))
          dim(fact); str(fact)
          corr_table <- round(cor(fact, use="pairwise.complete.obs"), digits=3); corr_table # similar to Fall
          factanal(na.omit(fact[,1:ncol(fact)]), 1, rotation="varimax") # factor loading is weaker than in fall 2014
          alpha(na.omit(d[,c("SS_A9_R_s15", "SS_A11_R_s15")])) # .59
          
      d$autonind_s15 <- apply(X = d[,c("SS_A9_R_s15","SS_A11_R_s15")], 
                        MARGIN =1, FUN = mean, na.rm = T)
      d <- moveme(d, c("SS_A9_R_s15", "SS_A11_R_s15", "autonind_s15"), "after", "SS_A12_s15")

###################################################################
###################### Fall 2015-2016 #############################
###################################################################
      d$SS_A9_R_f15 <- 6 - d$SS_A9_f15
      d$SS_A11_R_f15 <- 6 - d$SS_A11_f15
      fact <- subset(d, select = c(SS_A9_R_f15, SS_A10_f15, SS_A11_R_f15))
          dim(fact); str(fact)
          corr_table <- round(cor(fact, use="pairwise.complete.obs"), digits=3); corr_table # similar to Fall & Spring
          factanal(na.omit(fact[,1:ncol(fact)]), 1, rotation="varimax") # factor loading is strong like in fall 2014
          alpha(na.omit(d[,c("SS_A9_R_f15", "SS_A11_R_f15")])) # .61
          
      d$autonind_f15 <- apply(X = d[,c("SS_A9_R_f15","SS_A11_R_f15")], 
                        MARGIN =1, FUN = mean, na.rm = T)
      d <- moveme(d, c("SS_A9_R_f15", "SS_A11_R_f15", "autonind_f15"), "after", "SS_A12_f15")
      
###################################################################
###################### Spring 2015-2016 ###########################
###################################################################
      d$SS_A9_R_s16 <- 6 - d$SS_A9_s16
      d$SS_A11_R_s16 <- 6 - d$SS_A11_s16
      fact <- subset(d, select = c(SS_A9_R_s16, SS_A10_s16, SS_A11_R_s16))
          corr_table <- round(cor(fact, use="pairwise.complete.obs"), digits=3); corr_table # similar to Fall & Spring
          fact2 <- fact[complete.cases(fact),]
          eigen(cor(fact2))$values # only works with complete cases data. 
          factanal(na.omit(fact[,1:ncol(fact)]), 1, rotation="varimax")
          alpha(na.omit(d[,c("SS_A9_R_s16", "SS_A11_R_s16")])) # .67
          
      d$autonind_s16 <- apply(X = d[,c("SS_A9_R_s16","SS_A11_R_s16")], 
                        MARGIN =1, FUN = mean, na.rm = T)
      d <- moveme(d, c("SS_A9_R_s16", "SS_A11_R_s16", "autonind_s16"), "after", "SS_A12_s16")
```

```{r trust}
###################################################################
###################### Spring 2015-2016 ###########################
###################################################################

# create subset of variables to factor analyze
d$tr_control_R_s16 <- 7 - d$tr_control_s16

df_t <- subset(d, select = c("tr_best_s16", "tr_control_R_s16", "tr_fair_s16"))
 
    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot

# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 

# Alpha value for composite variable    
a <- alpha(na.omit(df_t[,c("tr_best_s16", "tr_control_R_s16", "tr_fair_s16")])) # .83 if middle item dropped

# Create composite variable
d$trust_cv_s16 <- apply(X = d[,c("tr_best_s16", "tr_fair_s16")], MARGIN=1, FUN=mean, na.rm=TRUE)

# Update variable order in main dataset
d <- moveme(d, c("trust_cv_s16"), "before", "tr_best_s16")
```

```{r stereotype threat}
# stereotype threat
      #ST1 = "In school, I worry that people will expect less of me because I have a learning difference."
      #ST2 = "I often worry that people will think I need help when I dont."
      #ST3 = "In school, I worry that people will expect me to make a mistake"
      #ST4 = "If I make a mistake in school, I worry about making people with learning differences look bad."

###################################################################
###################### Fall 2014-2015 #############################
###################################################################
##### only Fall 2014 asked 4-item scale for stereotype threat #####

    fact <- subset(d, select = c("ST1_f14", "ST2_f14", "ST3_f14", "ST4_f14"))
    dim(fact); str(fact)
    factanal(na.omit(fact[,1:ncol(fact)]), 1, rotation="varimax")
    alpha(na.omit(fact)) # .78
    d$ST_f14_4item <- apply(X = d[,c("ST1_f14", "ST2_f14", "ST3_f14", "ST4_f14")], 
                        MARGIN =1, FUN = mean, na.rm = T)
      hist(d$ST_f14_4item)
      
      # after Fall 2014 school only allows 2-item version for fear of insulting students. 
      # Fall 2014 ST
      alpha(na.omit(d[,c("ST2_f14", "ST3_f14")])) #.71
      d$ST_f14 <- apply(X = d[,c("ST2_f14", "ST3_f14")], 
                        MARGIN =1, FUN = mean, na.rm = T)
      hist(d$ST_f14)
      d <- moveme(d, c("ST_f14_4item", "ST_f14"), "after", "ST3_f14")

      # You didn't ask about stereotype threat in spring of 2015
          # SS_A12   How often do you worry that people will think less of you if you ask for help from a teacher?  is the closest question to ST in Spr15
          with(d, cor.test(SS_A12_s15, ST_f14)) # r = .23

###################################################################
###################### Spring 2014-2015 ###########################
###################################################################
          
# Was Stereotype threat measured this Spring? 
                      
###################################################################
###################### Fall 2015-2016 #############################
###################################################################
      alpha(na.omit(d[,c("ST2_f15", "ST3_f15")])) #.67
      d$ST_f15 <- apply(X = d[,c("ST2_f15", "ST3_f15")], 
                        MARGIN =1, FUN = mean, na.rm = T)
      hist(d$ST_f15)
      d <- moveme(d, "ST_f15", "after", "ST3_f15")

      # stereotype threat across the years
      d$ST_diff <- with(d, ST_f15 - ST_f14)
      d <- moveme(d, "ST_diff", "after", "ST_f15")

###################################################################
###################### Spring 2015-2016 ###########################
###################################################################
      
# create subset of variables to factor analyze
df_t <- subset(d, select = c('ST2_s16', 'ST3_s16'))
 
    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot

# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 

# Alpha value for composite variable    
a <- alpha(na.omit(df_t[,c("ST2_s16", "ST3_s16")])) # .61

# Create composite variable
d$st_cv_s16 <- apply(X = d[,c('ST2_s16', 'ST3_s16')], MARGIN=1, FUN=mean, na.rm=TRUE)

# Update variable order in main dataset
d <- moveme(d, c("st_cv_s16"), "before", "ST2_s16")
```

```{r belonging}
################################################################################
################################## Spring 2016 #################################
################################################################################

# measured in single item: Bel_s16
```

```{r learning goals}
################################################################################
################################## Spring 2016 #################################
################################################################################
      
# create subset of variables to factor analyze
df_t <- subset(d, select = c('learn_new_s16', 'mistak_s16'))
 
    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot

# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 

# Alpha value for composite variable    
a <- alpha(na.omit(df_t[,c('learn_new_s16', 'mistak_s16')])) # .75

# Create composite variable
d$learngoals_cv_s16 <- apply(X = d[,c('learn_new_s16', 'mistak_s16')], MARGIN=1, FUN=mean, na.rm=TRUE)

# Update variable order in main dataset
d <- moveme(d, c("learngoals_cv_s16"), "before", "learn_new_s16")
```

```{r strategic mindset strat_ms_cv}
names(d) [names(d) == 'SM1_r_s16'] <- 'SM1.r_s16'

d$SM1_R_s16 <- 7 - d$SM1.r_s16 # correctly recoded

fact <- subset(d, select = c(sm_better1_s16, 
                             sm_help_s16, 
                             sm_better2_s16, 
                             sm_stkhelp_s16, 
                             sm_bttrway_s16, 
                             sm_dobett_s16, 
                             SM1_R_s16, 
                             SM2_s16, 
                             SM_scav_s16)) 
            
        corr_table <- round(cor(fact, use="pairwise.complete.obs"), digits=3) #you need to determine how NAs should be treated. SPSS uses deleting NAs pairwise as a default, so R should do the same in order to compare the results

        corrplot(corr_table, method = "circle") #plot matrix
        
        scree_plot <- scree(corr_table,factors=TRUE,pc=TRUE,main="Scree plot",hline=NULL,add=FALSE); scree_plot 

# Begin Factor Analysis
dim(fact); str(fact)
factanal(na.omit(fact[,1:9]), 2, rotation="varimax")
    # SM1_R_s16 hangs with nothing
    # o/w classic SM hangs together and new items are seperate construct sans SM1_R_s16
bs_s <- alpha(na.omit(fact[,c(1:6)])) # .62 is magic number

# All components are 1-7 scored so standardization is not necessary before composite is computed: 
d$strat_ms_cv <- apply(X = d[,c('sm_better1_s16', 
                               'sm_help_s16', 
                               'sm_better2_s16', 
                               'sm_stkhelp_s16', 
                               'sm_bttrway_s16', 
                               'sm_dobett_s16')], 
                  MARGIN = 1, FUN = mean, na.rm = T)
      hist(d$strat_ms_cv)
```

The behaviors composite includes items like trying harder after difficulty, planning, considering alternatives, and taking advice.
```{r strategic and effortful behaviors}
################################################################################
################################## Spring 2016 #################################
################################################################################

# B_E1_s16    When something you are trying to learn is incredibly difficult, how often do you respond by trying harder?

      # B_E2_s16    When you fail to understand something, how often do you become discouraged and want to give up?
        d$B_E2_R_s16 <- 6 - d$B_E2_s16 # Reverse code items as needed
        d <- moveme(d, 'B_E2_R_s16', 'after', 'B_E2_s16')
  
# B_A1_s16    In school, how often do you work on improving your weaknesses rather than just working in areas where you are already strong?

# B_SP1_s16   Before you begin working on a project or assignment, how frequently do you consider different ways to prepare for it before choosing the best approach for you?

      # B_SP2_s16   When you are given a complicated school assignment, how frequently do you jump right in and start working without first spending time thinking about different ways to complete the assignment?
        d$B_SP2_R_s16 <- 6 - d$B_SP2_s16 # Reverse code items as needed
        d <- moveme(d, 'B_SP2_R_s16', 'after', 'B_SP2_s16')  
  
# B_SP3_s16   When youre trying to learn something new, how frequently do you start off by trying the approach your teacher recommends to see if it works best for you?

# B_SM4_s16   When youre doing schoolwork and get stuck, how frequently do you step back and consider a new approach?

# B_SM5_s16   When you struggle with something in school, how often do you try approaches that other students are using to see if they will work for you?  

# B_SM6_s16   When youre doing schoolwork and you get stuck, how often do you reach out to others to find out if they can suggest other ways of making progress?

      # B_SM7r_s16  When youre doing schoolwork and you get stuck, how often do you find it difficult to change your approach?
        d$B_SM7_R_s16 <- 6 - d$B_SM7r_s16 # Reverse code items as needed
        d <- moveme(d, 'B_SM7_R_s16', 'after', 'B_SM7r_s16')  

# create subset of variables to factor analyze
df_t <- subset(d, select = c(B_E1_s16:B_SM7_R_s16))
df_t <- subset(df_t, select = -c(B_E2_s16, B_SP2_s16, B_SM7r_s16)) # drop the original unreversed items. 

    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot
    corr_table %>% round(digits=2)
    
# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 
# Alpha value for composite variable    
alpha(na.omit(df_t[,c('B_E1_s16', 'B_A1_s16', 'B_SP1_s16', 'B_SP3_s16', 'B_SM4_s16', 'B_SM5_s16', 'B_SM6_s16')])) # .73

# Create composite variable
d$strat_beh_long_s16 <- apply(X = d[,c('B_E1_s16', 'B_A1_s16', 'B_SP1_s16', 'B_SP3_s16', 'B_SM4_s16', 'B_SM5_s16', 'B_SM6_s16')], MARGIN=1, FUN=mean, na.rm=TRUE)

# Update variable order in main dataset
d <- moveme(d, c("strat_beh_long_s16"), "before", "B_E1_s16")
```

### low alpha for small strat_beh scale
 - where did this strat_beh scale come from? 
```{r}
# create subset of variables to factor analyze
df_t <- subset(d, select = c("B_SP1_s16", "B_SP3_s16", "B_SM4_s16", 'SM2_s16', 'SM_scav_s16'))
 
    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot

# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 

# Alpha value for composite variable    
alpha(na.omit(df_t[,c("B_SP1_s16", "B_SP3_s16", "B_SM4_s16", 'SM2_s16', 'SM_scav_s16')])) 

# Create composite variable
d$strat_beh_small_s16 <- apply(X = d[,c("B_SP1_s16", "B_SP3_s16", "B_SM4_s16", 'SM2_s16', 'SM_scav_s16')], MARGIN=1, FUN=mean, na.rm=TRUE)

# Update variable order in main dataset
#d <- moveme(d, c("B_SP2_R_s16", "B_SM7_R_s16", "strat_beh_s16"), "after", "B_SM7r_s16")
```

```{r CV perceive teacher feedback as negative}
# Response Items: Never(1) -- Most of the Time (5)
# SS_B1  When your teacher gives you feedback, how frequently do you think carefully about it?
# SS_B2  After getting feedback from a teacher about how you could do something better, how frequently do you change how you do your schoolwork?
# SS_B3   How often do you disregard feedback from your teachers?
    # Correct directionality
    names(d) [names(d) == 'SS_B3_s16'] <- 'SS_B3r_s16'
    d$SS_B3_R_s16 <- 6 - d$SS_B3r_s16
# SS_F4  How often is feedback from your teachers really useful?
# SS_F7  How often do you feel like feedback from your teachers is mean?
# SS_A8  When youre struggling to learn something, how often do you seek out help?
# SS_A9  When you follow a teacher's suggestion for how to do something better, how often do you feel like youre being bossed around?
# SS_A10  When you use a teacher's suggestion for how to do something better in school, how often do you feel like youre taking charge of your education?
# SS_A11  When you use a teacher's suggestion for how you could do something better in school, how often do you feel like youre just doing what they're telling you to do?
# SS_A12   How often do you worry that people will think less of you if you ask for help from a teacher?  
  

# create subset of variables to factor analyze
df_t <- subset(d, select = c('SS_B1_s16', 'SS_B2_s16',  'SS_B3r_s16',  'SS_F4_s16',  'SS_F7_s16',  'SS_A8_s16',  'SS_A9_s16',  'SS_A10_s16',  'SS_A11_s16', 'SS_A12_s16'))
 
    # Save correlation table       
    corr_table <- df_t %>% 
    cor(use="pairwise.complete.obs") # SPSS deletes NAs pairwise as default
    
    corr_table %>% corrplot(method = "circle") # colored corr plot

# Scree Plot of Eigen values, set factors= in factanal() to number of eigen > 1.
corr_table %>% scree(factors=TRUE, pc=FALSE, main="Scree Plot", hline=NULL, add=FALSE) 
        
# Factor Analysis        
factanal(na.omit( df_t[, 1:ncol(df_t)] ), factors=1, rotation="varimax") 

# Alpha value for composite variable: hostile perception of teacher feedback    
alpha(na.omit(df_t[,c('SS_F7_s16', 'SS_A9_s16', 'SS_A11_s16')])) # .75

# Create composite variable
d$neg_tch_fb_s16 <- apply(X = d[,c('SS_F7_s16', 'SS_A9_s16', 'SS_A11_s16')], 
                      MARGIN=1, FUN=mean, na.rm=TRUE)

hist(d$neg_tch_fb_s16)

# Update variable order in main dataset
# d <- moveme(d, "neg_tch_fb_s16", "after", "SS_B1")
```

```{r grit 2014 only}
# GR2_f14 I often set a goal but later choose to pursue a different one. # Strongly Disagree (1) -- Strongly Agree (6)
    d$GR2_R_f14 <- 7 - d$GR2_f14
    d <- moveme(d, "GR2_R_f14", "after", "GR2_f14" )
# GR4_f14  I finish whatever I begin.

# Alpha value for composite variable    
a <- alpha(na.omit(d[,c("GR2_R_f14", "GR4_f14")])); a # .24

hist(d$GR2_R_f14)
hist(d$GR2_f14)

# Don't create composite variable, just use GR2_f14 for grit
```


## variable cleanup
```{r}
# rename variables
d$Pattern_1415 <- d$Pattern1415_1415; d$Pattern1415_1415 <- NULL                                       
d$GORT_date_1415 <- d$GORT_date1415_1415; d$GORT_date1415_1415 <- NULL                                                
d$Pattern1_1415 <- d$Pattern1_1415_1415; d$Pattern1_1415_1415 <- NULL                                                
d$Pattern2_1415 <- d$Pattern2_1415_1415; d$Pattern2_1415_1415 <- NULL                                                
d$Pattern3_1415 <- d$Pattern3_1415_1415; d$Pattern3_1415_1415 <- NULL                                                
d$Pattern4_1415 <- d$Pattern4_1415_1415; d$Pattern4_1415_1415 <- NULL                                                
d$Pattern5_1415 <- d$Pattern5_1415_1415; d$Pattern5_1415_1415 <- NULL                                                
d$Pattern6_1415 <- d$Pattern6_1415_1415; d$Pattern6_1415_1415 <- NULL                                                
d$Pattern7_1415 <- d$Pattern7_1415_1415; d$Pattern7_1415_1415 <- NULL                                                
d$Pattern8_1415 <- d$Pattern8_1415_1415; d$Pattern8_1415_1415 <- NULL                                                
d$Pattern9_1415 <- d$Pattern9_1415_1415; d$Pattern9_1415_1415 <- NULL 

d <- moveme(d, c("Pattern_1415", "GORT_date_1415", "Pattern1_1415", "Pattern2_1415", "Pattern3_1415", "Pattern4_1415", "Pattern5_1415", "Pattern6_1415", "Pattern7_1415", "Pattern8_1415", "Pattern9_1415"), "after", "GORT_RR_1415")

d <- moveme(d, c("ITI_amount_f14_R", "ITI_change_f14_R", "ITI_basic_f14_R", "ITI_comp_f14"), "after", "ITI_basic_f14")

d <- moveme(d, c("ITI_amount_s15_R", "ITI_change_s15_R", "ITI_basic_s15_R", "ITI_comp_s15"), "after", "ITI_basic_s15") 

d <- moveme(d, c("ITI_amount_f15_R", "ITI_change_f15_R", "ITI_basic_f15_R", "ITI_comp_f15"), "after", "ITI_basic_f15")                                                  

d$BRK_______________NEW_VARIABLES_______________BRK <- NA
d <- moveme(d, "BRK_______________NEW_VARIABLES_______________BRK", "after", "Yearbook_1516t1")

d <- moveme(d, "WISC_1415", "after", "WISC_TestDate_1415")                                                        
d <- moveme(d, "gort_1415", "after", "GORT_RR_1415")  

# drop meaningless variable
d <- subset(d, select = -c(STins_f14, idnum_ck1, SSins2_f14))

d[0,]
```


# XX. Save your merged data as .Rdata
```{r save you progress in an rdata file}
setwd("/Users/brozenkr/Box Sync/LD")
save.image(file = "03_LDMS_master.Rdata")
```
